### 2012-03-14 doc/icml2010stylefiles/example_paper.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%% ICML 2010 EXAMPLE LATEX SUBMISSION FILE %%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Use the following line _only_ if you're still using LaTeX 2.09.
%\documentstyle[icml2010,epsf,natbib]{article}
% If you rely on Latex2e packages, like most moden people use this:
\documentclass{article}
\usepackage[utf8]{inputenc}

% For figures
\usepackage{graphicx} % more modern
%\usepackage{epsfig} % less modern
\usepackage{subfigure}

% For citations
\usepackage{natbib}

% For algorithms
\usepackage{algorithm}
\usepackage{algorithmic}

% As of 2010, we use the hyperref package to produce hyperlinks in the
% resulting PDF.  If this breaks your system, please commend out the
% following usepackage line and replace \usepackage{icml2010} with
% \usepackage[nohyperref]{icml2010} above.
\usepackage{hyperref}

% Packages hyperref and algorithmic misbehave sometimes.  We can fix
% this with the following command.
\newcommand{\theHalgorithm}{\arabic{algorithm}}

% Employ the following version of the ``usepackage'' statement for
% submitting the draft version of the paper for review.  This will set
% the note in the first column to ``Under review.  Do not distribute.''
% \usepackage{icml2010} 
% Employ this version of the ``usepackage'' statement after the paper has
% been accepted, when creating the final version.  This will set the
% note in the first column to ``Appearing in''
\usepackage[accepted]{icml2010}

% The \icmltitle you define below is probably too long as a header.
% Therefore, a short form for the running title is supplied here:
\icmltitlerunning{Submission and Formatting Instructions for ICML 2010}

\begin{document}

\twocolumn[
\icmltitle{Number recognition for use in a Sudoku solver}

% It is OKAY to include author information, even for blind
% submissions: the style file will automatically remove it for you
% unless you've provided the [accepted] option to the icml2010
% package.
\icmlauthor{Ralf Vogler}{voglerr@in.tum.de}
\icmladdress{TU München}
\icmlauthor{Simon Grötzinger}{groetzin@in.tum.de}
\icmladdress{TU München}

% You may provide any keywords that you 
% find helpful for describing your paper; these are used to populate 
% the "keywords" metadata in the PDF but will not be shown in the document
\icmlkeywords{neural network, svm, logistic regression, synthetic data generation, machine learning}

\vskip 0.3in
]

\begin{abstract}
ICML 2010 full paper submissions are due February 1, 2010. Reviewing will
be blind to the identities of the authors, and therefore identifying
information should not appear in any way in papers submitted for
review. Submissions must be in PDF, 8 page length limit.
\end{abstract}

\section{Electronic Submission}
\label{submission}

As in the past few years, ICML 2010 will rely exclusively on
electronic formats for submission and review. We assume that all
authors will have access to standard software for word processing,
electronic mail, and web file transfer. Authors who do not have such
access should send email with their concerns to
\texttt{icml10.programchairs@gmail.com}.

\subsection{Templates for Papers}

Electronic templates for producing papers for submission are available
for \LaTeX\/ and Microsoft Word. Templates are accessible on the World
Wide Web at:\\
\textbf{\texttt{http://www.icml2010.org/}}

\noindent
Send questions about these electronic templates to
\texttt{icml10.programchairs@gmail.com}.

If you are using the templates, the formatting instructions below
will already be enforced. The content rules you must follow are:
\begin{itemize}
\item The maximum paper length is 8 pages.
\item Do not include author information or acknowledgments in your
  initial submission. 
\item Place figure captions {\em under} the figure (and omit titles from
  inside the graphic file itself).  Place table captions {\em over}
  the table.
\item References must include page numbers whenever possible and be as
  complete as possible.  Place multiple citations in chronological order.  
\end{itemize}
Please see below for details on each of these items.

\subsection{Submitting Papers}

Submission to ICML 2010 will be entirely electronic, via a web site
(not email).  The URL and information about the submission process
are available on the conference web site at

\textbf{\texttt{http://www.icml2010.org/sub.html}}

{\bf Paper Deadline:} The deadline for paper submission to ICML 2010
is Monday, February 1, 2010, at 11:59 p.m.~Samoa time.  If your full
submission does not reach us by this date, it will not be considered
for publication. There is no separate abstract submission this year.

{\bf Anonymous Submission:} To facilitate blind review, no identifying
author information should appear on the title page or in the paper
itself.  Section~\ref{author info} will explain the details of how to
format this.

{\bf Simultaneous Submission:} ICML will not accept any paper which,
at the time of submission, is under review for another conference or
has already been published. This policy also applies to papers that
overlap substantially in technical content with conference papers
under review or previously published. ICML submissions must not be
submitted to other conferences during ICML's review period. Authors
may submit to ICML substantially different versions of journal papers
that are currently under review by the journal, but not yet accepted
at the time of submission. Informal publications, such as technical
reports or papers in workshop proceedings which do not appear in
print, do not fall under these restrictions.

\medskip

To ensure our ability to print submissions, authors must provide their
manuscripts in \textbf{PDF} format.  Furthermore, please make sure
that files contain only Type-1 fonts (e.g.,~using the program {\tt
  pdffonts} in linux or using File/DocumentProperties/Fonts in
Acrobat).  Other fonts (like Type-3) might come from graphics files
imported into the document.

Authors using \textbf{Word} must convert their document to PDF.  Most
of the latest versions of Word have the facility to do this
automatically.  Submissions will not be accepted in Word format or any
format other than PDF. Really. We're not joking. Don't send Word.

Those who use \textbf{\LaTeX} to format their accepted papers need to
pay close attention to the typefaces used.  Specifically, when
producing the PDF by first converting the dvi output of \LaTeX\ to Postscript
the default behavior is to use non-scalable Type-3 PostScript bitmap
fonts to represent the standard \LaTeX\ fonts. The resulting document
is difficult to read in electronic form; the type appears fuzzy. To
avoid this problem, dvips must be instructed to use an alternative
font map.  This can be achieved with
something like the following commands:\\[0.5em]
{\bf dvips -Ppdf -tletter -G0 -o paper.ps paper.dvi}\\
{\bf ps2pdf paper.ps}\\[0.5em]
Note that it is a zero following the ``-G''.  This tells dvips to use
the config.pdf file (and this file refers to a better font mapping).

Another alternative is to use the \textbf{pdflatex} program instead of
straight \LaTeX. This program avoids the Type-3 font problem, however
you must ensure that all of the fonts are embedded (use {\tt
pdffonts}). If they are not, you need to configure pdflatex to use a
font map file that specifies that the fonts be embedded. Also you
should ensure that images are not downsampled or otherwise compressed
in a lossy way.

Note that the new 2010 style files use the {\tt hyperref} package to
make clickable links in documents.  If this causes problems for you,
add {\tt nohyperref} as one of the options to the {\tt icml2010}
usepackage statement.

\subsection{Reacting to Reviews}
We will continue the ICML tradition in which the authors are given the
option of providing a short reaction to the initial reviews. These
reactions will be taken into account in the discussion among the
reviewers and area chairs.

\subsection{Submitting Final Camera-Ready Copy}

The final versions of papers accepted for publication should follow the
same format and naming convention as initial submissions, except of
course that the normal author information (names and affiliations)
should be given.  See Section~\ref{final author} for details of how to
format this.

The footnote, ``Preliminary work.  Under review by the International
Conference on Machine Learning (ICML).  Do not distribute.'' must be
modified to ``Appearing in \textit{Proceedings of the
$\mathit{27}^{th}$ International Conference on Machine Learning},
Haifa, Israel, 2010.  Copyright 2010 by the author(s)/owner(s).''

For those using the \textbf{\LaTeX} style file, simply change
$\mathtt{\backslash usepackage\{icml2010\}}$ to

\verb|\usepackage[accepted]{icml2010}|

\noindent
Authors using \textbf{Word} must edit the
footnote on the first page of the document themselves.

Camera-ready copies should have the title of the paper as running head
on each page except the first one.  The running title consists of a
single line centered above a horizontal rule which is $1$ point thick.
The running head should be centered, bold and in $9$ point type.  The
rule should be $10$ points above the main text.  For those using the
\textbf{\LaTeX} style file, the original title is automatically set as running
head using the {\tt fancyhdr} package which is included in the ICML
2010 style file package.  In case that the original title exceeds the
size restrictions, a shorter form can be supplied by using

\verb|\icmltitlerunning{...}|

just before $\mathtt{\backslash begin\{document\}}$.
Authors using \textbf{Word} must edit the header of the document themselves.

\section{Format of the Paper}

All submissions must follow the same format to ensure the printer can
reproduce them without problems and to let readers more easily find
the information that they desire.

\subsection{Length and Dimensions}

Papers must not exceed eight (8) pages, including all figures, tables,
references, and appendices. Any submission that exceeds this page
limit or that diverges significantly from the format specified herein
will be rejected without review.

The text of the paper should be formatted in two columns, with an
overall width of 6.75 inches, height of 9.0 inches, and 0.25 inches
between the columns. The left margin should be 0.75 inches and the top
margin 1.0 inch (2.54~cm). The right and bottom margins will depend on
whether you print on US letter or A4 paper, but all final versions
must be produced for US letter size.

The paper body should be set in 10~point type with a vertical spacing
of 11~points. Please use Times Roman typeface throughout the text.

\subsection{Title}

The paper title should be set in 14~point bold type and centered
between two horizontal rules that are 1~point thick, with 1.0~inch
between the top rule and the top edge of the page. Capitalize the
first letter of content words and put the rest of the title in lower
case.

\subsection{Author Information for Submission}
\label{author info}

To facilitate blind review, author information must not appear.  If
you are using \LaTeX\/ and the \texttt{icml2010.sty} file, you may use
\verb+\icmlauthor{...}+ to specify authors.  The author information
will simply not be printed until {\tt accepted} is an argument to the
style file. Submissions that include the author information will not
be reviewed.

\subsubsection{Self-Citations}

If your are citing published papers for which you are an author, refer
to yourself in the third person. In particular, do not use phrases
that reveal your identity (e.g., ``in previous work \cite{langley00}, we 
have shown \ldots'').

Do not anonymize citations in the reference section by removing or
blacking out author names. The only exception are manuscripts that are
not yet published (e.g. under submission). If you choose to refer to
such unpublished manuscripts \cite{anonymous}, anonymized copies have to be submitted
as Supplementary Material via CMT. However, keep in mind that an ICML
paper should be self contained and should contain sufficient detail
for the reviewers to evaluate the work. In particular, reviewers are
not required to look a the Supplementary Material when writing their
review.

\subsubsection{Camera-Ready Author Information}
\label{final author}

If a paper is accepted, a final camera-ready copy must be prepared.
%
For camera-ready papers, author information should start 0.3~inches
below the bottom rule surrounding the title. The authors' names should
appear in 10~point bold type, electronic mail addresses in 10~point
small capitals, and physical addresses in ordinary 10~point type.
Each author's name should be flush left, whereas the email address
should be flush right on the same line. The author's physical address
should appear flush left on the ensuing line, on a single line if
possible. If successive authors have the same affiliation, then give
their physical address only once.

A sample file (in PDF) with author names is included in the ICML2010 style file package.

\subsection{Abstract}

The paper abstract should begin in the left column, 0.4~inches below
the final address. The heading `Abstract' should be centered, bold,
and in 11~point type. The abstract body should use 10~point type, with
a vertical spacing of 11~points, and should be indented 0.25~inches
more than normal on left-hand and right-hand margins. Insert
0.4~inches of blank space after the body. Keep your abstract brief,
limiting it to one paragraph and no more than six or seven sentences.

\subsection{Partitioning the Text}

You should organize your paper into sections and paragraphs to help
readers place a structure on the material and understand its
contributions.

\subsubsection{Sections and Subsections}

Section headings should be numbered, flush left, and set in 11~pt bold
type with the content words capitalized. Leave 0.25~inches of space
before the heading and 0.15~inches after the heading.

Similarly, subsection headings should be numbered, flush left, and set
in 10~pt bold type with the content words capitalized. Leave
0.2~inches of space before the heading and 0.13~inches afterward.

Finally, subsubsection headings should be numbered, flush left, and
set in 10~pt small caps with the content words capitalized. Leave
0.18~inches of space before the heading and 0.1~inches after the
heading.

Please use no more than three levels of headings.

\subsubsection{Paragraphs and Footnotes}

Within each section or subsection, you should further partition the
paper into paragraphs. Do not indent the first line of a given
paragraph, but insert a blank line between succeeding ones.

You can use footnotes\footnote{For the sake of readability, footnotes
should be complete sentences.} to provide readers with additional
information about a topic without interrupting the flow of the paper.
Indicate footnotes with a number in the text where the point is most
relevant. Place the footnote in 9~point type at the bottom of the
column in which it appears. Precede the first footnote in a column
with a horizontal rule of 0.8~inches.\footnote{Multiple footnotes can
appear in each column, in the same order as they appear in the text,
but spread them across columns and pages if possible.}

\begin{figure}[ht]
\vskip 0.2in
\begin{center}
\centerline{\includegraphics[width=\columnwidth]{icml_numpapers}}
\caption{Historical locations and number of accepted papers for International
  Machine Learning Conferences (ICML 1993 -- ICML 2008) and
  International Workshops on Machine Learning (ML 1988 -- ML
  1992). At the time this figure was produced, the number of
  accepted papers for ICML 2008 was unknown and instead estimated.}
\label{icml-historical}
\end{center}
\vskip -0.2in
\end{figure}

\subsection{Figures}

You may want to include figures in the paper to help readers visualize
your approach and your results. Such artwork should be centered,
legible, and separated from the text. Lines should be dark and at
least 0.5~points thick for purposes of reproduction, and text should
not appear on a gray background.

Label all distinct components of each figure. If the figure takes the
form of a graph, then give a name for each axis and include a legend
that briefly describes each curve. Do not include a title inside the
figure; instead, the caption should serve this function.

Number figures sequentially, placing the figure number and caption
{\it after\/} the graphics, with at least 0.1~inches of space before
the caption and 0.1~inches after it, as in
Figure~\ref{icml-historical}.  The figure caption should be set in
9~point type and centered unless it runs two or more lines, in which
case it should be flush left.  You may float figures to the top or
bottom of a column, and you may set wide figures across both columns
(use the environment {\tt figure*} in \LaTeX), but always place
two-column figures at the top or bottom of the page.

\subsection{Algorithms}

If you are using \LaTeX, please use the ``algorithm'' and ``algorithmic'' 
environments to format pseudocode. These require 
the corresponding stylefiles, algorithm.sty and 
algorithmic.sty, which are supplied with this package. 
Algorithm~\ref{alg:example} shows an example.

\begin{algorithm}[tb]
   \caption{Bubble Sort}
   \label{alg:example}
\begin{algorithmic}
   \STATE {\bfseries Input:} data $x_i$, size $m$
   \REPEAT
   \STATE Initialize $noChange = true$.
   \FOR{$i=1$ {\bfseries to} $m-1$}
   \IF{$x_i > x_{i+1}$} 
   \STATE Swap $x_i$ and $x_{i+1}$
   \STATE $noChange = false$
   \ENDIF
   \ENDFOR
   \UNTIL{$noChange$ is $true$}
\end{algorithmic}
\end{algorithm}

\subsection{Tables}

You may also want to include tables that summarize material. Like 
figures, these should be centered, legible, and numbered consecutively. 
However, place the title {\it above\/} the table with at least 
0.1~inches of space before the title and the same after it, as in 
Table~\ref{sample-table}. The table title should be set in 9~point 
type and centered unless it runs two or more lines, in which case it
should be flush left.

% Note use of \abovespace and \belowspace to get reasonable spacing 
% above and below tabular lines.

\begin{table}[t]
\caption{Classification accuracies for naive Bayes and flexible 
Bayes on various data sets.}
\label{sample-table}
\vskip 0.15in
\begin{center}
\begin{small}
\begin{sc}
\begin{tabular}{lcccr}
\hline
\abovespace\belowspace
Data set & Naive & Flexible & Better? \\
\hline
\abovespace
Breast    & 95.9$\pm$ 0.2& 96.7$\pm$ 0.2& $\surd$ \\
Cleveland & 83.3$\pm$ 0.6& 80.0$\pm$ 0.6& $\times$\\
Glass2    & 61.9$\pm$ 1.4& 83.8$\pm$ 0.7& $\surd$ \\
Credit    & 74.8$\pm$ 0.5& 78.3$\pm$ 0.6&         \\
Horse     & 73.3$\pm$ 0.9& 69.7$\pm$ 1.0& $\times$\\
Meta      & 67.1$\pm$ 0.6& 76.5$\pm$ 0.5& $\surd$ \\
Pima      & 75.1$\pm$ 0.6& 73.9$\pm$ 0.5&         \\
\belowspace
Vehicle   & 44.9$\pm$ 0.6& 61.5$\pm$ 0.4& $\surd$ \\
\hline
\end{tabular}
\end{sc}
\end{small}
\end{center}
\vskip -0.1in
\end{table}

Tables contain textual material that can be typeset, as contrasted 
with figures, which contain graphical material that must be drawn. 
Specify the contents of each row and column in the table's topmost
row. Again, you may float tables to a column's top or bottom, and set
wide tables across both columns, but place two-column tables at the
top or bottom of the page.

\subsection{Citations and References}

Please use APA reference format regardless of your formatter
or word processor. If you rely on the \LaTeX\/ bibliographic 
facility, use {\tt natbib.sty} and {\tt icml2010.bst} 
included in the style-file package to obtain this format.

Citations within the text should include the authors' last names and
year. If the authors' names are included in the sentence, place only
the year in parentheses, for example when referencing Arthur Samuel's
pioneering work \yrcite{Samuel59}. Otherwise place the entire
reference in parentheses with the authors and year separated by a
comma \cite{Samuel59}. List multiple references separated by
semicolons \cite{kearns89,Samuel59,mitchell80}. Use the `et~al.'
construct only for citations with three or more authors or after
listing all authors to a publication in an earlier reference \cite{MachineLearningI}.

Authors should cite their own work in the third person
in the initial version of their paper submitted for blind review.
Please refer to Section~\ref{author info} for detailed instructions on how to
cite your own papers.

Use an unnumbered first-level section heading for the references, and 
use a hanging indent style, with the first line of the reference flush
against the left margin and subsequent lines indented by 10 points. 
The references at the end of this document give examples for journal
articles \cite{Samuel59}, conference publications \cite{langley00}, book chapters \cite{Newell81}, books \cite{DudaHart2nd}, edited volumes \cite{MachineLearningI}, 
technical reports \cite{mitchell80}, and dissertations \cite{kearns89}.

Alphabetize references by the surnames of the first authors, with
single author entries preceding multiple author entries. Order
references for the same authors by year of publication, with the
earliest first. Make sure that each reference includes all relevant
information (e.g., page numbers).

\subsection{Software and Data}

We strongly encourage the publication of software and data with the
camera-ready version of the paper whenever appropriate.  This can be
done by including a URL in the camera-ready copy.  However, do not
include URLs that reveal your institution or identity in your
submission for review.  Instead, provide an anonymous URL or upload
the material as ``Supplementary Material'' into the CMT reviewing
system.  Note that reviewers are not required to look a this material
when writing their review.

% Acknowledgements should only appear in the accepted version. 
\section*{Acknowledgments}

\textbf{Do not} include acknowledgements in the initial version of
the paper submitted for blind review.

If a paper is accepted, the final camera-ready version can (and
probably should) include acknowledgements. In this case, please
place such acknowledgements in an unnumbered section at the
end of the paper. Typically, this will include thanks to reviewers
who gave useful comments, to colleagues who contributed to the ideas, 
and to funding agencies and corporate sponsors that provided financial 
support.

% In the unusual situation where you want a paper to appear in the
% references without citing it in the main text, use \nocite
\nocite{langley00}

\bibliography{example_paper}
\bibliographystyle{icml2010}

\end{document}

% This document was modified from the file originally made available by
% Pat Langley and Andrea Danyluk for ICML-2K. This 2010 version was
% created by Thorsten Joachims & Johannes Fuernkranz, 
% slightly modified from the 2009 version by Kiri Wagstaff and 
% Sam Roweis's 2008 version, which is slightly modified from 
% Prasad Tadepalli's 2007 version which is a lightly 
% changed version of the previous year's version by Andrew Moore, 
% which was in turn edited from those of Kristian Kersting and 
% Codrina Lauth. Alex Smola contributed to the algorithmic style files.

### 2012-03-14 nn/config.m
num_labels fixed, saving data file with -mat-binary

@@ -7,16 +7,13 @@ function config
 
 	input_layer_size  = imgSize(data_folder);  % equals resolution of images
 	hidden_layer_size = 25;   % 25 hidden units
-	num_labels = 10;          % 10 labels, from 1 to 10 (label 0)
+	num_labels = 9;           % 9 labels, from 1 to 9
 	
 	set_size = 100;			  % size of the k sets
 	
-	% libsvm_path = "./";
-	% addpath(libsvm_path);
-	
 	use_nn = true;
-	use_svm = true;
-	use_lr = true;
+	use_svm = false;
+	use_lr = false;
 	
 	save("config.mat");
 endfunction

### 2012-03-14 nn/main.m

@@ -12,8 +12,8 @@ if exist(cache)
 else
     fprintf('Loading generated images...\n')
     [X,y] = processImages(data_folder);
-    save("-binary", cache, "X", "y");
-    % save(cache, "X", "y");
+    save("-mat-binary", cache, "X", "y");
+    %save(cache, "X", "y");
 endif
 X(:,size(X,2)+1) = y;
 

### 2012-03-14 nn/train.m

@@ -4,11 +4,10 @@ function [Theta1, Theta2] = train(X, y)
 	m = size(X, 1);
 
 	% Randomly display 10 samples
-	fprintf('Show 10 random samples\n')
-	sel = randperm(size(X, 1));
-	sel = sel(1:10);
-
-	%displayData(X(sel, :));
+	% fprintf('Show 10 random samples\n')
+	% sel = randperm(size(X, 1));
+	% sel = sel(1:10);
+	% displayData(X(sel, :));
 
 	% Randomly initialize weights
 	fprintf('Randomly initializing Neural Network Parameters...\n')

### 2012-03-14 nn.py
added python nn file and changed config file format to be matlab compatible

@@ -1,0 +1,32 @@
+#from libs.pybrain.structure import FeedForwardNetwork
+import os, sys
+import scipy.io
+
+#current_dir = os.path.dirname(os.path.abspath(__file__))
+#sys.path.append(os.path.join(current_dir, 'libs'))
+
+from pybrain.structure import FeedForwardNetwork, LinearLayer, SigmoidLayer, FullConnection
+
+def main():
+    mat = scipy.io.loadmat('data/cache.mat')
+    X = mat['X']
+    #print X[0]
+    y = mat['y']
+    config = scipy.io.loadmat('nn/config.mat')
+    n = FeedForwardNetwork()
+    inLayer = LinearLayer(config['input_layer_size'][0][0])
+    hiddenLayer = SigmoidLayer(config['hidden_layer_size'][0][0])
+    outLayer = LinearLayer(config['num_labels'][0][0])
+    n.addInputModule(inLayer)
+    n.addModule(hiddenLayer)
+    n.addOutputModule(outLayer)
+    in_to_hidden = FullConnection(inLayer, hiddenLayer)
+    hidden_to_out = FullConnection(hiddenLayer, outLayer)
+    n.addConnection(in_to_hidden)
+    n.addConnection(hidden_to_out)
+    n.sortModules()
+    print n
+    print n.activate(X[0])
+    
+
+main()

### 2012-03-14 nn/config.m

@@ -15,5 +15,5 @@ function config
 	use_svm = false;
 	use_lr = false;
 	
-	save("config.mat");
+	save("-mat-binary", "config.mat");
 endfunction

### 2012-03-14 gen-data.py
Neural Network with pyBrain
images less random

@@ -6,10 +6,10 @@ folder = 'data'
 if os.path.exists(folder):
     shutil.rmtree(folder)
 
-fonts = ['arial.ttf', 'ariali.ttf', 'comicbd.ttf', 'HARNGTON.TTF', 'FREESCPT.TTF'] #'LHANDW.TTF', 'HoboStd.otf']
+fonts = ['arial.ttf', 'ariali.ttf', 'comicbd.ttf']#, 'HARNGTON.TTF', 'FREESCPT.TTF'] #'LHANDW.TTF', 'HoboStd.otf']
 size_mu = 19
 size_sig = 3
-pos_sig = 2
+pos_sig = 0
 color_sig = 100
 color_bg = 'white'
 rotate_sig = 25

### 2012-03-14 nn.py

@@ -1,32 +1,62 @@
 #from libs.pybrain.structure import FeedForwardNetwork
 import os, sys
-import scipy.io
+from scipy.io import loadmat
 
 #current_dir = os.path.dirname(os.path.abspath(__file__))
 #sys.path.append(os.path.join(current_dir, 'libs'))
 
-from pybrain.structure import FeedForwardNetwork, LinearLayer, SigmoidLayer, FullConnection
+from pybrain.tools.shortcuts     import buildNetwork
+from pybrain.structure           import SigmoidLayer, TanhLayer, SoftmaxLayer
+from pybrain.datasets            import ClassificationDataSet
+from pybrain.supervised.trainers import BackpropTrainer
+from pybrain.utilities           import percentError
+from pybrain.tests.helpers import gradientCheck
+
+n_in = 400
+n_hidden = 25
+n_classes = 9
 
 def main():
-    mat = scipy.io.loadmat('data/cache.mat')
+    config = loadmat('nn/config.mat')
+    cache = config['data_folder'][0].replace('../', '')+'/'+config['cache_file'][0]
+    if not os.path.exists(cache):
+        raise Exception('Cache file ('+cache+') is missing and has to be generated first! Run octave version...')
+    mat = loadmat(cache)
     X = mat['X']
-    #print X[0]
     y = mat['y']
-    config = scipy.io.loadmat('nn/config.mat')
-    n = FeedForwardNetwork()
-    inLayer = LinearLayer(config['input_layer_size'][0][0])
-    hiddenLayer = SigmoidLayer(config['hidden_layer_size'][0][0])
-    outLayer = LinearLayer(config['num_labels'][0][0])
-    n.addInputModule(inLayer)
-    n.addModule(hiddenLayer)
-    n.addOutputModule(outLayer)
-    in_to_hidden = FullConnection(inLayer, hiddenLayer)
-    hidden_to_out = FullConnection(hiddenLayer, outLayer)
-    n.addConnection(in_to_hidden)
-    n.addConnection(hidden_to_out)
-    n.sortModules()
-    print n
-    print n.activate(X[0])
     
+    ds = ClassificationDataSet(n_in, 1, nb_classes=n_classes)
+    for (x, y) in zip(X, y):
+        ds.addSample(x, [y-1])
+    # ds.assignClasses()
+    
+    tstdata, trndata = ds.splitWithProportion(0.25)
[…]

### 2012-03-14 nn/config.m

@@ -9,7 +9,7 @@ function config
 	hidden_layer_size = 25;   % 25 hidden units
 	num_labels = 9;           % 9 labels, from 1 to 9
 	
-	set_size = 100;			  % size of the k sets
+	set_size = 1125;		  % size of the k sets
 	
 	use_nn = true;
 	use_svm = false;

### 2012-03-14 ML-Sudoku/Main.cpp
CV: cleanup: removed gl & glut

@@ -2,51 +2,33 @@
 #include <iostream>
 #include <iomanip>
 
-#include <glut.h>
-
 #include <cv.h>
 #include <highgui.h>
 
 #include <math.h>
 
-#include "PoseEstimation.h"
-
 using namespace std;
 
-// Added in Exercise 9 - Start *****************************************************************
-#define PI 3.14159265
-
-struct Position { double x,y,z; };
-
 bool debugmode = false;
-bool balldebug = false;
-// Added in Exercise 9 - End *****************************************************************
+bool adaptiveTh = true;
+
+const int MIN = 120;
+const int S = 360;
 
 int thresh = 100;
-CvCapture* cap;
-
-int bw_thresh = 40;
-
-CvMemStorage* memStorage;
-
-// Added in Exercise 9 - Start *****************************************************************
-float resultMatrix_005A[16];
-float resultMatrix_0272[16];
-float resultTransposedMatrix[16];
-float snowmanLookVector[4];
-int towards = 0x005A;
-int towardsList[2] = {0x005a, 0x0272};
-int towardscounter = 0;
-Position ballpos;
-int ballspeed = 100;
-// Added in Exercise 9 - End *****************************************************************
-
+int bw_thresh = 80;
 //camera settings
 const int width = 640; 
 const int height = 480;
-const int camangle = 35;
 
-unsigned char bkgnd[width*height*3];
+CvCapture* cap;
+CvMemStorage* memStorage;
+
+float resultMatrix_005A[16];
+float resultMatrix_0272[16];
+float resultTransposedMatrix[16];
+
+#define PI 3.14159265
 
 void trackbarHandler(int pos) {
 	thresh = pos;
@@ -61,7 +43,7 @@ void initVideoStream() {
 
 	if (!cap) {
 		cout << "No webcam found, using video file\n";
-		cap = cvCaptureFromFile("C:\\Andi\\Uni\\SVNs\\far_intern\\teaching\\2010WS\\AR\\Exercises\\Solutions\\MarkerMovie.mpg");
+		cap = cvCaptureFromFile("C:\\MarkerMovie.mpg");
 		if (!cap) {
 			cout << "No video file found. Exiting.\n";
[…]

### 2012-03-14 nn.py
- nn with pybrain
- overview diagram
- results/parameters from training are saved in models.mat
- predict uses the models.mat file to predict outcome

@@ -6,27 +6,52 @@ import scipy.io
 #sys.path.append(os.path.join(current_dir, 'libs'))
 
 from pybrain.structure import FeedForwardNetwork, LinearLayer, SigmoidLayer, FullConnection
+from pybrain.tools.shortcuts import buildNetwork
+from pybrain.datasets import ClassificationDataSet
+from pybrain.supervised.trainers import BackpropTrainer
+from pybrain.utilities import percentError
+from pybrain.tools.validation import CrossValidator
 
-def main():
-    mat = scipy.io.loadmat('data/cache.mat')
-    X = mat['X']
-    #print X[0]
-    y = mat['y']
-    config = scipy.io.loadmat('nn/config.mat')
-    n = FeedForwardNetwork()
-    inLayer = LinearLayer(config['input_layer_size'][0][0])
-    hiddenLayer = SigmoidLayer(config['hidden_layer_size'][0][0])
-    outLayer = LinearLayer(config['num_labels'][0][0])
-    n.addInputModule(inLayer)
-    n.addModule(hiddenLayer)
-    n.addOutputModule(outLayer)
-    in_to_hidden = FullConnection(inLayer, hiddenLayer)
-    hidden_to_out = FullConnection(hiddenLayer, outLayer)
-    n.addConnection(in_to_hidden)
-    n.addConnection(hidden_to_out)
-    n.sortModules()
-    print n
-    print n.activate(X[0])
+
+config = scipy.io.loadmat('nn/config.mat')
+input_layer_size = config['input_layer_size'][0][0]
+hidden_layer_size = config['hidden_layer_size'][0][0]
+num_labels = int(config['num_labels'][0][0])
+mat = scipy.io.loadmat('data/cache.mat')
+X = mat['X']
+y = mat['y']
+
+def prepare_dataset():
+    alldata = ClassificationDataSet(input_layer_size, 1, nb_classes=num_labels)
+    for (i,j) in zip(X,y):
+        alldata.addSample(i,j-1)
+    return alldata
     
+def get_2layer_nn():
+    return buildNetwork(input_layer_size, hidden_layer_size, num_labels)
+    
+def main():
+    alldata = prepare_dataset()
+
+    #trainer = BackpropTrainer(n, alldata)
+    #alldata._convertToOneOfMany()
+    #CrossValidator(trainer, 
+    
+    tstdata, trndata = alldata.splitWithProportion(0.25)
+    trndata._convertToOneOfMany()
[…]

### 2012-03-14 nn/config.m

@@ -2,18 +2,20 @@ function config
 	% Setup parameters
 	data_folder = '../data';  % folder that contains the generated images
     cache_file  = 'cache.mat';% name of the cache file
+	models_file = 'models.mat'; % for saving trained nn/svm/lr models/parameters
 	lambda = 1;               % Regularization
-	maxIter = 200;            % number of max. iterations
+	maxIter = 50;            % number of max. iterations
 
 	input_layer_size  = imgSize(data_folder);  % equals resolution of images
 	hidden_layer_size = 25;   % 25 hidden units
 	num_labels = 9;           % 9 labels, from 1 to 9
 	
-	set_size = 100;			  % size of the k sets
+	set_size = 1125;			  % size of the k sets
+	do_cross_validation = false;	% if set to false we train only with one partition of train/validation sets
 	
 	use_nn = true;
-	use_svm = false;
-	use_lr = false;
+	use_svm = true;
+	use_lr = true;
 	
 	save("-mat-binary", "config.mat");
 endfunction

### 2012-03-14 nn/predict.m

@@ -1,20 +1,21 @@
-function p = predict(Theta1, Theta2, X)
-%PREDICT Predict the label of an input given a trained neural network
-%   p = PREDICT(Theta1, Theta2, X) outputs the predicted label of X given the
-%   trained weights of a neural network (Theta1, Theta2)
-
-% Useful values
-m = size(X, 1);
-num_labels = size(Theta2, 1);
-
-% You need to return the following variables correctly 
-p = zeros(size(X, 1), 1);
-
-h1 = sigmoid([ones(m, 1) X] * Theta1');
-h2 = sigmoid([ones(m, 1) h1] * Theta2');
-[dummy, p] = max(h2, [], 2);
-
-% =========================================================================
-
-
+function [pNN, pSVM, pLR] = predict(X)
+	load('config.mat');
+	load(models_file);
+	
+	m = size(X, 1);
+	
+	if (exist("Theta1") && exist("Theta2"))
+		% predict with NN
+		pNN = predictNN(Theta1, Theta2, X);
+	endif
+	if (exist("model"))
+		% predict with SVM		
+		%[predict_label, acc_svm, dec_values] = svmpredict(y_val,X_val,model);
+		printf("...ignore this output, it is from libsvm...");
+		[pSVM,_,_] = svmpredict(ones(m,1),X,model);
+	endif
+	if (exist("Theta"))
+		pLR = predictLogReg(Theta,X);
+	endif
+	printf("predicted values:\n  NN: %d\n  SVM: %d\n  LR: %d\n", pNN, pSVM, pLR);
 end

### 2012-03-14 nn/predictLogReg.m

@@ -10,10 +10,10 @@ num_labels = size(Theta, 2);
 p = zeros(size(X, 1), 1);
 
 % Add ones to the X data matrix
-X = [ones(m, 1) X];
+%X = [ones(m, 1) X];
 
 % predict and choose max element in row
-p = sigmoid(X*Theta);
+p = sigmoid([ones(m, 1) X]*Theta);
 [_,p] = max(p,[],2);
 
 

### 2012-03-14 nn/predictNN.m

@@ -1,0 +1,17 @@
+function p = predictNN(Theta1, Theta2, X)
+%PREDICT Predict the label of an input given a trained neural network
+%   p = PREDICT(Theta1, Theta2, X) outputs the predicted label of X given the
+%   trained weights of a neural network (Theta1, Theta2)
+
+% Useful values
+m = size(X, 1);
+num_labels = size(Theta2, 1);
+
+% You need to return the following variables correctly 
+p = zeros(size(X, 1), 1);
+
+h1 = sigmoid([ones(m, 1) X] * Theta1');
+h2 = sigmoid([ones(m, 1) h1] * Theta2');
+[dummy, p] = max(h2, [], 2);
+
+end

### 2012-03-14 nn/config.m
fixed config

@@ -3,13 +3,14 @@ function config
 	data_folder = '../data';  % folder that contains the generated images
     cache_file  = 'cache.mat';% name of the cache file
 	lambda = 1;               % Regularization
-	maxIter = 200;            % number of max. iterations
+	maxIter = 100;            % number of max. iterations
 
 	input_layer_size  = imgSize(data_folder);  % equals resolution of images
 	hidden_layer_size = 25;   % 25 hidden units
 	num_labels = 9;           % 9 labels, from 1 to 9
 	
 	set_size = 1125;		  % size of the k sets
+	do_cross_validation = false; % if set to false we train only one time (no cross validation)
 	
 	use_nn = true;
 	use_svm = false;

### 2012-03-14 doc/icml2010stylefiles/paper.tex
copied latex template
- python nn uses values from config.m
- added some text and overview image to docu

--- 2012-03-14 doc/icml2010stylefiles/example_paper.tex
@@
 % Therefore, a short form for the running title is supplied here:
-\icmltitlerunning{Submission and Formatting Instructions for ICML 2010}
+\icmltitlerunning{Number recognition for use in a Sudoku solver}
 
@@
 \end{abstract}
+
+\section{Introduction}
+Number and letter recognition is key in many real-world applications.
+Digitalizing print media, automated sorting of letters, or recognition of signs in robotics or driver assistance systems require well performing algorithms for character recognition.
+In the following we analyze the performance of different techniques for number recognition, namely Neural Networks, Support Vector Machines and Logistic Regression, by using them in an automated sudoku solver.
+
+\section{Overview}
+% trim=l b r t
+\centerline{\includegraphics[width=\columnwidth, trim=25 10 25 10]{overview}}
+
+blabla
+
+\section{Generation of Training Data}
+\subsection{Image Generation}
+\subsection{Image Processing}
+
+\section{Training}
+\subsection{Neural Network}
+\subsection{Support Vector Machine}
+\subsection{Logistic Regrssion}
+
+\section{Image Recognition and Preprocessing}
+
+\section{Number Prediction}
+
+\section{Evaluation}
 
@@
 
-\bibliography{example_paper}
+\bibliography{paper}
 \bibliographystyle{icml2010}

### 2012-03-14 nn.py

@@ -12,22 +12,25 @@ from pybrain.supervised.trainers import BackpropTrainer
 from pybrain.utilities           import percentError
 from pybrain.tests.helpers import gradientCheck
 
-n_in = 400
-n_hidden = 25
-n_classes = 9
+#n_in = 400
+#n_hidden = 25
+#n_classes = 9
+config = loadmat('nn/config.mat')
+cache = config['data_folder'][0].replace('../', '')+'/'+config['cache_file'][0]
+if not os.path.exists(cache):
+    raise Exception('Cache file ('+cache+') is missing and has to be generated first! Run octave version...')
+mat = loadmat(cache)
+X = mat['X']
+y = mat['y']
+n_in = int(config['input_layer_size'][0][0])
+n_hidden = int(config['hidden_layer_size'][0][0])
+n_classes = int(config['num_labels'][0][0])
+weightdecay = config['lambda'][0][0]
 
-def main():
-    config = loadmat('nn/config.mat')
-    cache = config['data_folder'][0].replace('../', '')+'/'+config['cache_file'][0]
-    if not os.path.exists(cache):
-        raise Exception('Cache file ('+cache+') is missing and has to be generated first! Run octave version...')
-    mat = loadmat(cache)
-    X = mat['X']
-    y = mat['y']
-    
+def main():    
     ds = ClassificationDataSet(n_in, 1, nb_classes=n_classes)
-    for (x, y) in zip(X, y):
-        ds.addSample(x, [y-1])
+    for (x, l) in zip(X, y):
+        ds.addSample(x, l-1)
     # ds.assignClasses()
     
     tstdata, trndata = ds.splitWithProportion(0.25)
@@ -45,9 +48,9 @@ def main():
     # print trndata['input'][0], trndata['target'][0], trndata['class'][0]
     
     net = buildNetwork(trndata.indim, n_hidden, trndata.outdim, bias=True, hiddenclass=SigmoidLayer, outclass=SoftmaxLayer) #, recurrent=True)
-    trainer = BackpropTrainer(net, trndata)
+    trainer = BackpropTrainer(net, trndata, learningrate=0.01, weightdecay=weightdecay)
 
-    for i in xrange(50):
+    for i in xrange(20):
         # print trainer.train()
         # train the network for 1 epoch
         trainer.trainEpochs(1)

### 2012-03-14 nn/config.m

@@ -2,12 +2,13 @@ function config
 	% Setup parameters
 	data_folder = '../data';  % folder that contains the generated images
     cache_file  = 'cache.mat';% name of the cache file
-	lambda = 1;               % Regularization
+	models_file = 'models.mat'; % contains trained models/parameters of nn/svm/lr
+	lambda = 0;               % Regularization
 	maxIter = 100;            % number of max. iterations
 
 	input_layer_size  = imgSize(data_folder);  % equals resolution of images
 	hidden_layer_size = 25;   % 25 hidden units
-	num_labels = 9;           % 9 labels, from 1 to 9
+	num_labels = 9;           % numbers 1-9
 	
 	set_size = 1125;		  % size of the k sets
 	do_cross_validation = false; % if set to false we train only one time (no cross validation)

### 2012-03-14 nn/predict.m

@@ -1,4 +1,7 @@
 function [pNN, pSVM, pLR] = predict(X)
+	pNN = -1;
+	pSVM = -1;
+	pLR = -1;
 	load('config.mat');
 	load(models_file);
 	
@@ -17,5 +20,5 @@ function [pNN, pSVM, pLR] = predict(X)
 	if (exist("Theta"))
 		pLR = predictLogReg(Theta,X);
 	endif
-	printf("predicted values:\n  NN: %d\n  SVM: %d\n  LR: %d\n", pNN, pSVM, pLR);
+	printf("predicted values (-1 means that model was not trained):\n  NN: %d\n  SVM: %d\n  LR: %d\n", pNN, pSVM, pLR);
 end

### 2012-03-14 nn.py
- lambda and maxIter can be passed to config function
- nn.py uses maxIter from config

@@ -26,6 +26,7 @@ n_in = int(config['input_layer_size'][0][0])
 n_hidden = int(config['hidden_layer_size'][0][0])
 n_classes = int(config['num_labels'][0][0])
 weightdecay = config['lambda'][0][0]
+maxIter = int(config['maxIter'][0][0])
 
 def main():    
     ds = ClassificationDataSet(n_in, 1, nb_classes=n_classes)
@@ -48,9 +49,10 @@ def main():
     # print trndata['input'][0], trndata['target'][0], trndata['class'][0]
     
     net = buildNetwork(trndata.indim, n_hidden, trndata.outdim, bias=True, hiddenclass=SigmoidLayer, outclass=SoftmaxLayer) #, recurrent=True)
+    print "Learning with lambda = %f. Performing %i iterations" % (weightdecay, maxIter)
     trainer = BackpropTrainer(net, trndata, learningrate=0.01, weightdecay=weightdecay)
 
-    for i in xrange(20):
+    for i in xrange(maxIter):
         # print trainer.train()
         # train the network for 1 epoch
         trainer.trainEpochs(1)

### 2012-03-14 nn/config.m

@@ -1,10 +1,10 @@
-function config
+function config (lambda=1, maxIter=100)
 	% Setup parameters
 	data_folder = '../data';  % folder that contains the generated images
     cache_file  = 'cache.mat';% name of the cache file
 	models_file = 'models.mat'; % contains trained models/parameters of nn/svm/lr
-	lambda = 0;               % Regularization
-	maxIter = 100;            % number of max. iterations
+	%lambda = 1;               % Regularization
+	%maxIter = 100;            % number of max. iterations
 
 	input_layer_size  = imgSize(data_folder);  % equals resolution of images
 	hidden_layer_size = 25;   % 25 hidden units

### 2012-03-14 split-image.py
Python script to split image into fields
> seed fill in corners to eliminate borders

@@ -1,0 +1,77 @@
+import Image, ImageDraw, ImageFont, ImageStat
+import random, os, shutil
+
+input = 'cv/sudoku.png'
+out = 'input/'
+tmp = out+'tmp/'
+size = 20
+m = 0
+threshold = 240 # brighter fields are empty
+
+if not os.path.exists(input):
+    print input+' is missing! Capture image using cv.exe first.'
+    exit()
+
+if os.path.exists(out):
+    shutil.rmtree(out)
+os.makedirs(tmp)
+
+def within(img, x, y):
+    w, h = img.size
+    return x >= 0 and x < w and y >= 0 and y < h
+
+BORDER_COLOR = 240
+def flood_fill(img, x, y, value):
+    "Flood fill on a region of non-BORDER_COLOR pixels."
+    p = img.load()
+    if not within(img, x, y) or p[x, y] >= BORDER_COLOR:
+        return
+    edge = [(x, y)]
+    p[x, y] = value
+    while edge:
+        newedge = []
+        for (x, y) in edge:
+            for (s, t) in ((x+1, y), (x-1, y), (x, y+1), (x, y-1)):
+                if within(img, s, t) and p[s, t] < BORDER_COLOR: #not in (BORDER_COLOR, value):
+                    # print s, t, p[s, t], value
+                    p[s, t] = value
+                    newedge.append((s, t))
+        edge = newedge
+
+def process(img):
+    "flood fill white in every corner to remove unwanted edges"
+    # better: http://packages.python.org/mahotas
+    o = 0
+    s = size-o-1
+    flood_fill(img, o, o, 255)
+    flood_fill(img, s, o, 255)
+    flood_fill(img, o, s, 255)
+    flood_fill(img, s, s, 255)
+    return img
+
+def brightness(img):
+   im = img.convert('L')
+   stat = ImageStat.Stat(im)
+   return stat.mean[0]
+
+img = Image.open(input)
+img.thumbnail((9*size, 9*size), Image.ANTIALIAS)
+img.save(tmp+'resized.png', 'PNG')
+histogram = []
+for i in range(9*9):
+    x = i%9*size
+    y = i/9*size
+    cropped = img.crop((x+m, y+m, x+size-m, y+size-m))
+    cropped = process(cropped)
+    file = str(i+1)+'.png'
+    cropped.save(tmp+file, 'PNG')
+    
+    b = brightness(cropped)
+    histogram.append(b)
+    if b > threshold:
[…]

### 2012-03-14 split-image.py
run-solver

@@ -59,8 +59,10 @@ img.thumbnail((9*size, 9*size), Image.ANTIALIAS)
 img.save(tmp+'resized.png', 'PNG')
 histogram = []
 for i in range(9*9):
-    x = i%9*size
-    y = i/9*size
+    row = i%9
+    col = i/9
+    x = row*size
+    y = col*size
     cropped = img.crop((x+m, y+m, x+size-m, y+size-m))
     cropped = process(cropped)
     file = str(i+1)+'.png'

### 2012-03-14 nn/config.m
- octave file for processing sudoku input images and predict values
- python script for handling input processing, prediction, and solving (todo)
-

@@ -3,6 +3,7 @@ function config (lambda=1, maxIter=100)
 	data_folder = '../data';  % folder that contains the generated images
     cache_file  = 'cache.mat';% name of the cache file
 	models_file = 'models.mat'; % contains trained models/parameters of nn/svm/lr
+	input_folder = '../input'; % folder that contains all images from the observed sudoku
 	%lambda = 1;               % Regularization
 	%maxIter = 100;            % number of max. iterations
 

### 2012-03-14 nn/predict.m

@@ -1,11 +1,11 @@
 function [pNN, pSVM, pLR] = predict(X)
-	pNN = -1;
-	pSVM = -1;
-	pLR = -1;
+	m = size(X, 1);
+	pNN = ones(m,1)*(-1);
+	pSVM = ones(m,1)*(-1);
+	pLR = ones(m,1)*(-1);
 	load('config.mat');
 	load(models_file);
 	
-	m = size(X, 1);
 	
 	if (exist("Theta1") && exist("Theta2"))
 		% predict with NN
@@ -20,5 +20,8 @@ function [pNN, pSVM, pLR] = predict(X)
 	if (exist("Theta"))
 		pLR = predictLogReg(Theta,X);
 	endif
-	printf("predicted values (-1 means that model was not trained):\n  NN: %d\n  SVM: %d\n  LR: %d\n", pNN, pSVM, pLR);
+	pNN
+	pSVM
+	pLR
+	%printf("predicted values (-1 means that model was not trained):\n  NN: %d\n  SVM: %d\n  LR: %d\n", pNN, pSVM, pLR);
 end

### 2012-03-14 nn/processInput.m

@@ -1,0 +1,19 @@
+load('config.mat');
+% set path
+dataPath = sprintf("./%s/", input_folder);
+path = regexprep(dataPath, '\\', '/');
+
+X = []; % matrix to store the image vectors
+
+files = readdir(path);
+for x=1:length(files)
+	if(!isdir(files{x}))
+		currentFile = sprintf("%s%s", path, files{x});
+		rgbImg = imread(currentFile);
+		grayImg = rgb2gray(rgbImg);
+		grayImgVector = double(grayImg(:)');
+		X(size(X,1)+1,:) = grayImgVector/255;
+	endif
+end
+[pNN, pSVM, pLR] = predict(X);
+save("-mat-binary", "inputAndPredictedValues.mat", "X", "pNN", "pSVM", "pLR");

### 2012-03-14 processAndPredict.py

@@ -1,0 +1,17 @@
+#from libs.pybrain.structure import FeedForwardNetwork
+import os, sys
+from scipy.io import loadmat
+
+print "processing input images with octave and writing file nn/input.mat"
+os.system('run-processInput.bat')
+
+
+input = loadmat('nn/inputAndPredictedValues.mat')
+
+# input matrix of detected sudoku numbers
+X = input['X']
+
+# predicted values from octave
+pNN = input['pNN']
+pSVM = input['pSVM']
+pLR = input['pLR']

### 2012-03-14 nn/processInput.m
save field vector when processing input

@@ -4,6 +4,7 @@ dataPath = sprintf("./%s/", input_folder);
 path = regexprep(dataPath, '\\', '/');
 
 X = []; % matrix to store the image vectors
+field = []; % vector that stores the corresponding sudoku field of a row in X matrix
 
 files = readdir(path);
 for x=1:length(files)
@@ -13,7 +14,9 @@ for x=1:length(files)
 		grayImg = rgb2gray(rgbImg);
 		grayImgVector = double(grayImg(:)');
 		X(size(X,1)+1,:) = grayImgVector/255;
+		field(size(field,1)+1,:) = str2num(strsplit(files{x},"."){1});
 	endif
 end
 [pNN, pSVM, pLR] = predict(X);
-save("-mat-binary", "inputAndPredictedValues.mat", "X", "pNN", "pSVM", "pLR");
+field
+save("-mat-binary", "inputAndPredictedValues.mat", "X", "field", "pNN", "pSVM", "pLR");

### 2012-03-14 processAndPredict.py
renamed nn to octave, again ;)

@@ -2,16 +2,22 @@
 import os, sys
 from scipy.io import loadmat
 
-print "processing input images with octave and writing file nn/input.mat"
+print "processing input images with octave and writing file octave/inputAndPredictedValues.mat"
 os.system('run-processInput.bat')
 
 
-input = loadmat('nn/inputAndPredictedValues.mat')
+input = loadmat('octave/inputAndPredictedValues.mat')
 
 # input matrix of detected sudoku numbers
 X = input['X']
 
 # predicted values from octave
+field = input['field']
 pNN = input['pNN']
 pSVM = input['pSVM']
 pLR = input['pLR']
+
+classes = zip(field, pNN)
+print classes
+for i,c in classes:
+    print i/9, i%9, c

### 2012-03-15 nn.py
fixes

@@ -14,7 +14,7 @@ from pybrain.tests.helpers       import gradientCheck
 #n_in = 400
 #n_hidden = 25
 #n_classes = 9
-config = loadmat('nn/config.mat')
+config = loadmat('octave/config.mat')
 cache = config['data_folder'][0].replace('../', '')+'/'+config['cache_file'][0]
 if not os.path.exists(cache):
     raise Exception('Cache file ('+cache+') is missing and has to be generated first! Run octave version...')
@@ -26,6 +26,8 @@ n_hidden = int(config['hidden_layer_size'][0][0])
 n_classes = int(config['num_labels'][0][0])
 weightdecay = config['lambda'][0][0]
 maxIter = int(config['maxIter'][0][0])
+
+from pybrain.tools.customxml import NetworkWriter
 
 def main():    
     ds = ClassificationDataSet(n_in, 1, nb_classes=n_classes)
@@ -62,5 +64,7 @@ def main():
               "  train error: %5.2f%%" % trnresult, \
               "  test error: %5.2f%%" % tstresult
     # print trainer.trainUntilConvergence()
+
+    NetworkWriter.writeToFile(net, 'net')
     
 main()

### 2012-03-15 octave/main.m

@@ -1,7 +1,9 @@
 % Initialization
 clear ; close all; clc
 
-config();
+if !exist('config.mat')
+	config();
+endif
 load('config.mat');
 
 % Load images

### 2012-03-15 octave/processInput.m
predict classes of processed fields

@@ -1,20 +1,20 @@
 load('config.mat');
 % set path
 dataPath = sprintf("./%s/", input_folder);
-path = regexprep(dataPath, '\\', '/');
+path = regexprep(dataPath, '/', '\\');
 
 X = []; % matrix to store the image vectors
 field = []; % vector that stores the corresponding sudoku field of a row in X matrix
 
-files = readdir(path);
+files = glob(sprintf('%s%s', path, '*.png'));
 for x=1:length(files)
 	if(!isdir(files{x}))
-		currentFile = sprintf("%s%s", path, files{x});
-		rgbImg = imread(currentFile);
+		rgbImg = imread(files{x});
 		grayImg = rgb2gray(rgbImg);
 		grayImgVector = double(grayImg(:)');
 		X(size(X,1)+1,:) = grayImgVector/255;
-		field(size(field,1)+1,:) = str2num(strsplit(files{x},"."){1});
+        [dir, name, ext, ver] = fileparts(files{x});
+		field(size(field,1)+1,:) = str2num(name);
 	endif
 end
 [pNN, pSVM, pLR] = predict(X);

### 2012-03-15 gen-data.py
gen-data.py: switch for random/concrete

@@ -2,17 +2,26 @@ import Image, ImageDraw, ImageFont
 import random, os, shutil
 
 folder = 'data'
+concrete = True
 
 if os.path.exists(folder):
     shutil.rmtree(folder)
 
-fonts = ['arial.ttf', 'ariali.ttf', 'comicbd.ttf']#, 'HARNGTON.TTF', 'FREESCPT.TTF'] #'LHANDW.TTF', 'HoboStd.otf']
-size_mu = 19
-size_sig = 3
-pos_sig = 0
-color_sig = 100
+if(concrete):
+    fonts = ['arialbd.ttf', 'ariblk.ttf']
+    size_mu = 18
+    size_sig = 1
+    pos_sig = 1
+    color_sig = 50
+    rotate_sig = 5
+else:
+    fonts = ['arial.ttf', 'ariali.ttf', 'comicbd.ttf']#, 'HARNGTON.TTF', 'FREESCPT.TTF'] #'LHANDW.TTF', 'HoboStd.otf']
+    size_mu = 19
+    size_sig = 3
+    pos_sig = 0
+    color_sig = 100
+    rotate_sig = 25
 color_bg = 'white'
-rotate_sig = 25
 nums = xrange(1,10)
 samples = xrange(0,500)
 width, height = 20, 20

### 2012-03-15 nn-predict.py
added nn-predict to predict values based on a net trained with pybrain

@@ -1,0 +1,14 @@
+from pybrain.tools.customxml import NetworkReader
+from scipy.io import loadmat
+
+def main():
+    net = NetworkReader.readFrom('net')
+    input = loadmat('octave/inputAndPredictedValues.mat')
+    X = input['X']
+    field = input['field']
+    for i in range(0,len(X)):
+        r = net.activate(X[i])
+        #print field[i][0]
+        print "%d - %d" %(field[i][0], r.argmax()+1)
+        
+main()

### 2012-03-15 nn.py

@@ -65,6 +65,6 @@ def main():
               "  test error: %5.2f%%" % tstresult
     # print trainer.trainUntilConvergence()
 
-    NetworkWriter.writeToFile(net, 'net')
+    NetworkWriter.writeToFile(net, 'net.xml')
     
 main()

### 2012-03-15 nn-predict.py
fix

@@ -2,7 +2,7 @@ from pybrain.tools.customxml import NetworkReader
 from scipy.io import loadmat
 
 def main():
-    net = NetworkReader.readFrom('net')
+    net = NetworkReader.readFrom('net.xml')
     input = loadmat('octave/inputAndPredictedValues.mat')
     X = input['X']
     field = input['field']

### 2012-03-15 gen-data.py
trying to improve image processing and data-gen

@@ -8,12 +8,12 @@ if os.path.exists(folder):
     shutil.rmtree(folder)
 
 if(concrete):
-    fonts = ['arialbd.ttf', 'ariblk.ttf']
-    size_mu = 18
+    fonts = ['arialbd.ttf', 'comicbd.ttf'] #, 'arialbd.ttf', 'ariblk.ttf']
+    size_mu = 16
     size_sig = 1
     pos_sig = 1
-    color_sig = 50
-    rotate_sig = 5
+    color_sig = 10
+    rotate_sig = 0
 else:
     fonts = ['arial.ttf', 'ariali.ttf', 'comicbd.ttf']#, 'HARNGTON.TTF', 'FREESCPT.TTF'] #'LHANDW.TTF', 'HoboStd.otf']
     size_mu = 19

### 2012-03-15 split-image.py

@@ -6,7 +6,7 @@ out = 'input/'
 tmp = out+'tmp/'
 size = 20
 m = 0
-threshold = 240 # brighter fields are empty
+threshold = 250 # brighter fields are empty
 
 if not os.path.exists(input):
     print input+' is missing! Capture image using cv.exe first.'
@@ -24,7 +24,7 @@ BORDER_COLOR = 240
 def flood_fill(img, x, y, value):
     "Flood fill on a region of non-BORDER_COLOR pixels."
     p = img.load()
-    if not within(img, x, y) or p[x, y] >= BORDER_COLOR:
+    if not within(img, x, y) or p[x, y] >= 240:
         return
     edge = [(x, y)]
     p[x, y] = value
@@ -32,27 +32,40 @@ def flood_fill(img, x, y, value):
         newedge = []
         for (x, y) in edge:
             for (s, t) in ((x+1, y), (x-1, y), (x, y+1), (x, y-1)):
-                if within(img, s, t) and p[s, t] < BORDER_COLOR: #not in (BORDER_COLOR, value):
-                    # print s, t, p[s, t], value
-                    p[s, t] = value
-                    newedge.append((s, t))
+                if within(img, s, t) and p[s, t] < BORDER_COLOR: #abs(org - p[s, t]) < 50: #not in (BORDER_COLOR, value):
+                        # print s, t, p[s, t], value
+                        p[s, t] = value
+                        newedge.append((s, t))
         edge = newedge
 
+FILL = 255
 def process(img):
     "flood fill white in every corner to remove unwanted edges"
     # better: http://packages.python.org/mahotas
-    o = 0
-    s = size-o-1
-    flood_fill(img, o, o, 255)
-    flood_fill(img, s, o, 255)
-    flood_fill(img, o, s, 255)
-    flood_fill(img, s, s, 255)
+    
+    for o in range(2): # fill from corners going inwards with an offset o
+        s = size-o-1
+        flood_fill(img, o, o, FILL)
+        flood_fill(img, s, o, FILL)
+        flood_fill(img, o, s, FILL)
+        flood_fill(img, s, s, FILL)
+
+    for o in range(1): # go around all edges
+        for i in range(size):
+            flood_fill(img, o, i, FILL)
+            flood_fill(img, i, size-o, FILL)
+            flood_fill(img, size-o, i, FILL)
+            flood_fill(img, i, o, FILL)
+
[…]

### 2012-03-15 octave/config.m
- added examples to doc
- added nn to doc
- added comment in config

@@ -1,4 +1,9 @@
-function config (lambda=1, maxIter=100)
+% this file contains global settings for python/octave scripts
+% ========================= CAUTION =========================
+% if you change any value here, you have to call config()
+% within octave to regenerate the config.mat file
+
+function config (lambda=0.01, maxIter=100)
 	% Setup parameters
 	data_folder = '../data';  % folder that contains the generated images
     cache_file  = 'cache.mat';% name of the cache file
@@ -14,8 +19,8 @@ function config (lambda=1, maxIter=100)
 	set_size = 1125;		  % size of the k sets
 	do_cross_validation = false; % if set to false we train only one time (no cross validation)
 	
-	use_nn = true;
-	use_svm = false;
+	use_nn = false;
+	use_svm = true;
 	use_lr = false;
 	
 	save("-mat-binary", "config.mat");

### 2012-03-15 octave/processInput.m

@@ -18,5 +18,4 @@ for x=1:length(files)
 	endif
 end
 [pNN, pSVM, pLR] = predict(X);
-field
 save("-mat-binary", "inputAndPredictedValues.mat", "X", "field", "pNN", "pSVM", "pLR");

### 2012-03-15 gen-data.py
use full resolution of captured image instead of scaling down immediately

@@ -11,7 +11,7 @@ if(concrete):
     fonts = ['arialbd.ttf', 'comicbd.ttf'] #, 'arialbd.ttf', 'ariblk.ttf']
     size_mu = 16
     size_sig = 1
-    pos_sig = 1
+    pos_sig = 3
     color_sig = 10
     rotate_sig = 0
 else:

### 2012-03-15 processAndPredict.py
validation with concrete Sudoku

@@ -17,6 +17,9 @@ pNN = input['pNN']
 pSVM = input['pSVM']
 pLR = input['pLR']
 
+def section(s):
+    print 
+    print '-'*6+' '+s+' '+'-'*6
 
 def unique(seq):
     seen = set()
@@ -26,8 +29,10 @@ def unique(seq):
 field_param = ''
 rows = [[] for i in range(9)]
 cols = [[] for i in range(9)]
-for i,c in zip(field, pNN):
+predicted_values = [0 for i in range(9*9)]
+for i,c in zip(field, pSVM):
     i, c = int(i), int(c)
+    predicted_values[i-1] = c
     row = int((i-1)/9)
     col = int((i-1)%9)
     print i, row, col, c
@@ -35,6 +40,27 @@ for i,c in zip(field, pNN):
     rows[row].append(c)
     cols[col].append(c)
 
+# accuracy
+section('Accuracy')
+# filter(lambda a: a 0= 0, predicted_values)
+solution = [
+3,0,0, 2,4,0, 0,6,0,
+0,4,0, 0,0,0, 0,5,3,
+1,8,9, 6,3,5, 4,0,0,
+0,0,0, 0,8,0, 2,0,0,
+0,0,7, 4,9,6, 8,0,1,
+8,9,3, 1,5,0, 6,0,4,
+0,0,1, 9,2,0, 5,0,0,
+2,0,0, 3,0,0, 7,4,0,
+9,6,0, 5,0,0, 3,0,2]
+correct = [a for (a,b) in zip(solution,predicted_values) if a==b]
+failed  = [(i+1,a,b) for (i,a,b) in zip(range(9*9), solution,predicted_values) if a!=b]
+print 'Correct: ', len(correct), 'Failed: ', len(failed)
+print failed
+print 'Accuracy: ', len(correct)/(9.*9)
+
+
+section('Valid Sudoku?')
 error = False
 for row, col in zip(rows, cols):
     if row != unique(row) or col != unique(col):
@@ -45,4 +71,6 @@ for row, col in zip(rows, cols):
 if error:
         print 'WARNING: Solver will run forever (just used to display Sudoku field)'
         # exit()
+
+section('Solver')
 os.system('java -cp . Sudoku'+field_param)

### 2012-03-15 gen-data.py
parameter testing

@@ -8,9 +8,9 @@ if os.path.exists(folder):
     shutil.rmtree(folder)
 
 if(concrete):
-    fonts = ['arialbd.ttf', 'comicbd.ttf'] #, 'arialbd.ttf', 'ariblk.ttf']
+    fonts = ['arialbd.ttf', 'comicbd.ttf', 'ariblk.ttf'] #, 'arialbd.ttf', 'ariblk.ttf']
     size_mu = 16
-    size_sig = 1
+    size_sig = 2
     pos_sig = 3
     color_sig = 10
     rotate_sig = 0

### 2012-03-15 octave/config.m

@@ -3,7 +3,7 @@
 % if you change any value here, you have to call config()
 % within octave to regenerate the config.mat file
 
-function config (lambda=0.01, maxIter=100)
+function config (lambda=1, maxIter=100)
 	% Setup parameters
 	data_folder = '../data';  % folder that contains the generated images
     cache_file  = 'cache.mat';% name of the cache file

### 2012-03-15 octave/main.m

@@ -65,7 +65,7 @@ for i = 0:loops
 		printf('\n=== svm in iteration %i\n', i);
 		% svm
 		t=cputime;
-		model = svmtrain(y_train, X_train, sprintf("-q -c %i -t 2", lambda));
+		model = svmtrain(y_train, X_train, sprintf("-q -c %f -t 2 -d 10", lambda)); % default d=3
 		cpu_time_svm += cputime-t;
 		%printf('Total cpu time for training: %f seconds\n', cputime-t);
 		fprintf('Test Set Accuracy (SVM): ');

### 2012-03-15 octave/config.m
added gamma parameter and gridSearch

@@ -3,7 +3,7 @@
 % if you change any value here, you have to call config()
 % within octave to regenerate the config.mat file
 
-function config (lambda=0.01, maxIter=100)
+function config (lambda=0.01, maxIter=100, gamma=1)
 	% Setup parameters
 	data_folder = '../data';  % folder that contains the generated images
     cache_file  = 'cache.mat';% name of the cache file

### 2012-03-15 octave/gridSearchSVM.m

@@ -1,0 +1,13 @@
+% Initialization
+%clear ; close all; clc
+
+C = [2^-5,  2^-3,  2^-1,  2^1,  2^3,  2^5,  2^7,  2^9,  2^11, 2^13];
+g = [2^-15, 2^-13, 2^-11, 2^-9, 2^-7, 2^-5, 2^-3, 2^-1, 2^1,  2^3];
+
+for cIt=1:length(C)
+	for gIt=1:length(g)
+		config(C(cIt), 100, g(gIt));
+		printf("\n=====> next paramater pair: C = %f, g = %f\n", C(cIt), g(gIt));
+		main
+	end
+end

### 2012-03-15 octave/main.m

@@ -1,5 +1,5 @@
 % Initialization
-clear ; close all; clc
+%clear ; close all; clc
 
 if !exist('config.mat')
 	config();
@@ -28,7 +28,7 @@ average_accuracy_log_reg = 0;
 cpu_time_nn = 0;
 cpu_time_svm = 0;
 cpu_time_lr = 0;
-number_of_sets = ceil(length(y)/set_size)
+number_of_sets = ceil(length(y)/set_size);
 
 if (do_cross_validation)
 	loops = number_of_sets-1;
@@ -37,7 +37,9 @@ else
 endif
 
 for i = 0:loops
-	printf('\n===== Starting iteration %i (next training/test set combination) =====\n', i);
+	if (do_cross_validation)
+		printf('\n===== Starting iteration %i (next training/test set combination) =====\n', i);
+	endif
 	b = min((i+1)*set_size, size(y)); % needed if last set is smaller
 	sel = all_indexed_mixed(i*set_size+1 : b); % selecting the current set of row indexes
 	
@@ -62,10 +64,14 @@ for i = 0:loops
 	endif
 	
 	if (use_svm)
-		printf('\n=== svm in iteration %i\n', i);
+		if (do_cross_validation)
+			printf('\n=== svm in iteration %i\n', i);
+		endif
 		% svm
 		t=cputime;
-		model = svmtrain(y_train, X_train, sprintf("-q -c %i -t 2", lambda));
+		lambda
+		gamma
+		model = svmtrain(y_train, X_train, sprintf("-q -c %f -t 2 -g %f", lambda, gamma));
 		cpu_time_svm += cputime-t;
 		%printf('Total cpu time for training: %f seconds\n', cputime-t);
 		fprintf('Test Set Accuracy (SVM): ');

### 2012-03-15 octave/main.m
fixed gamma parameter

@@ -65,7 +65,7 @@ for i = 0:loops
 		printf('\n=== svm in iteration %i\n', i);
 		% svm
 		t=cputime;
-		model = svmtrain(y_train, X_train, sprintf("-q -c %f -t 2 -d 10", lambda)); % default d=3
+		model = svmtrain(y_train, X_train, sprintf("-q -c %f -t 2 -d 10 -g %f", lambda, gamma)); % default d=3
 		cpu_time_svm += cputime-t;
 		%printf('Total cpu time for training: %f seconds\n', cputime-t);
 		fprintf('Test Set Accuracy (SVM): ');

### 2012-03-15 octave/main.m
predict using PyBrain

@@ -65,7 +65,7 @@ for i = 0:loops
 		printf('\n=== svm in iteration %i\n', i);
 		% svm
 		t=cputime;
-		model = svmtrain(y_train, X_train, sprintf("-q -c %f -t 2 -d 10 -g %f", lambda, gamma)); % default d=3
+		model = svmtrain(y_train, X_train, sprintf("-q -c %f -t 2 -g %f", lambda, gamma));
 		cpu_time_svm += cputime-t;
 		%printf('Total cpu time for training: %f seconds\n', cputime-t);
 		fprintf('Test Set Accuracy (SVM): ');

### 2012-03-15 doc/icml2010stylefiles/paper.tex
moved overview diagram to doc folder
pdf removed from gitignore because of includes for latex

@@
 Digitalizing print media, automated sorting of letters, or recognition of signs in robotics or driver assistance systems require well performing algorithms for character recognition.
-In the following we analyze the performance of different techniques for number recognition, namely Neural Networks, Support Vector Machines and Logistic Regression, by using them in an automated sudoku solver.
+In the following we analyze the performance of different techniques for number recognition, namely Neural Networks, Support Vector Machines and Logistic Regression, by using them in an automated Sudoku solver.
 
@@
 % trim=l b r t
-\centerline{\includegraphics[width=\columnwidth, trim=25 10 25 10]{overview}}
-
-blabla
+\begin{figure}[ht]
+	\centerline{\includegraphics[width=\columnwidth, trim=25 30 25 30]{overview}}
+	\caption{Overview}
+	\label{fig:overview}
+\end{figure}
 
@@
 \subsection{Image Generation}
+\begin{figure}[ht]
+	\centerline{\includegraphics[scale=0.25]{examples}}
+	\caption{Examples of generated numbers}
+	\label{fig:generated_numbers}
+\end{figure}
 \subsection{Image Processing}
@@
 \subsection{Neural Network}
+\begin{figure}[ht]
+	\centerline{\includegraphics[width=\columnwidth, trim=0 30 0 30]{nn}}
+	\caption{Neural Network with one hidden layer}
+	\label{fig:nn}
+\end{figure}
 \subsection{Support Vector Machine}
@@
 \section{Evaluation}
+
+\section{Future Work}
+\color{red} include information if a sudoku is unsolvable \color{black}
 

### 2012-03-15 doc/icml2010stylefiles/paper.tex
worked on svm part in doc

@@
 
+% custom imports
+\usepackage{amsmath}
+
 \begin{document}
@@
 Digitalizing print media, automated sorting of letters, or recognition of signs in robotics or driver assistance systems require well performing algorithms for character recognition.
-In the following we analyze the performance of different techniques for number recognition, namely Neural Networks, Support Vector Machines and Logistic Regression, by using them in an automated Sudoku solver.
+In the following we analyze the performance of different techniques for number recognition, namely Neural Networks, Support Vector Machines and Logistic Regression, by using them in an automated sudoku solver.
 
@@
 \end{figure}
+
 \subsection{Image Processing}
@@
 \end{figure}
-\subsection{Support Vector Machine}
-\subsection{Logistic Regrssion}
+\subsection{Support Vector Machines}
+The next technique we applied to the discussed problem are Support Vector Machines \cite{Bishop200608} which try to solve the following optimization problem:
+\begin{align*}
+\min_{w,b} &~~ \frac{1}{2}||w||^2 + C \sum_{i=1}^{n} \xi_i\\
+\text{subject to } &~~ y_i(w^T \phi(x_i) + b) - 1 + \xi_i \ge 0\\
+\text{and} &~~ \xi_i \ge 0.
+\end{align*}
+Using a radial basis function (RBF) as kernel we performed a grid search over $C$ and $\gamma$ to find those parameters that provide the highest accuracy on the validation set.
+We used the open source library ``libsvm'' \cite{chang2011libsvm} wrapped by some octave scripts for setting up data sets and performing the grid search.
+\subsection{Logistic Regression}
 
@@
 \color{red} include information if a sudoku is unsolvable \color{black}
+
+\section{References}
+
+\bibliography{paper}
+\bibliographystyle{icml2010}
 

### 2012-03-15 doc/icml2010stylefiles/paper.tex

@@
 \end{figure}
-
 \subsection{Image Processing}

### 2012-03-15 doc/icml2010stylefiles/paper.tex
abstract

--- 2012-03-15 doc/icml2010stylefiles/paper.tex
@@
 \begin{abstract}
-ICML 2010 full paper submissions are due February 1, 2010. Reviewing will
-be blind to the identities of the authors, and therefore identifying
-information should not appear in any way in papers submitted for
-review. Submissions must be in PDF, 8 page length limit.
+The goal of this work is to implement number recognition for a Soduko solver.
+First we will describe the generation of synthetic data, in this case labelled images of numbers.
+Then we compare different machine learning techniques for supervised learning using the generated data.
+Once the training is complete, images of real Sudokus are captured using the webcam.
+The different methods then try to correctly predict each number.
 \end{abstract}

### 2012-03-15 ML-Sudoku/Main.cpp
run-cv, capture and exit on 'c'

@@ -349,7 +349,7 @@ void idle()
 
 				if ( fabs(c) < 0.001 ) //lines parallel?
 				{
-					std::cout << "lines parallel" << std::endl;
+					//std::cout << "lines parallel" << std::endl;
 					continue;
 				}
 
@@ -432,9 +432,9 @@ void idle()
 	if (key == 27) { run = false; }
 	else if (key == 100) debug = !debug;
 	else if (key == 99){
-		cout << "caputure!\n";
+		cout << "Image caputured!\n";
 		cvSaveImage("sudoku.png", iplMarker);
-		//run = false;
+		run = false;
 	}
 
 	isFirstStripe = true;
@@ -455,12 +455,12 @@ void cleanup()
 	cvDestroyWindow ("Converted Image");
 	cvDestroyWindow ("Stripe");
 	cvDestroyWindow ("Marker");
-	cout << "Finished\n";
+	cout << "Finished!\n";
 }
 
 int main(int argc, char* argv[]) 
 {
-	cout << "Press 'Esc' exit\n";
+	cout << "Press 'Esc' to exit\n";
 	cout << "Press 'c' to capture once the image seems right (yellow dot in left upper corner)...\n";
 	// setup OpenCV
 	init();

### 2012-03-15 octave/main.m
output import config in main and added debug imshow in predict

@@ -5,6 +5,20 @@ if !exist('config.mat')
 	config();
 endif
 load('config.mat');
+lambda, maxIter, gamma
+use_nn, use_svm, use_lr
+if exist(models_file)
+    load(models_file);
+endif
+if (exist("Theta1") && exist("Theta2"))
+    disp("nn exists");
+endif
+if (exist("model"))
+    disp("svm exists");
+endif
+if (exist("Theta"))
+    disp("lr exists");
+endif
 
 % Load images
 cache = sprintf("%s/%s", data_folder, cache_file);

### 2012-03-15 processAndPredict.py
refactored: different prediction methods with accuracy-checks
params

@@ -3,6 +3,16 @@ import os, sys
 from nn_predict import predict
 from scipy.io import loadmat
 
+def section(s):
+    print 
+    print 
+    print '-'*6+' '+s+' '+'-'*6
+
+def unique(seq):
+    seen = set()
+    seen_add = seen.add
+    return [ x for x in seq if x not in seen and not seen_add(x)]
+
 print "processing input images with octave and writing file octave/inputAndPredictedValues.mat"
 os.system('run-processInput.bat')
 
@@ -12,40 +22,6 @@ input = loadmat('octave/inputAndPredictedValues.mat')
 # input matrix of detected sudoku numbers
 X = input['X']
 
-# predicted values from octave
-field = input['field']
-pNN = input['pNN']
-pSVM = input['pSVM']
-pLR = input['pLR']
-pPyBrain = predict()
-p = pSVM
-
-def section(s):
-    print 
-    print '-'*6+' '+s+' '+'-'*6
-
-def unique(seq):
-    seen = set()
-    seen_add = seen.add
-    return [ x for x in seq if x not in seen and not seen_add(x)]
-
-field_param = ''
-rows = [[] for i in range(9)]
-cols = [[] for i in range(9)]
-predicted_values = [0 for i in range(9*9)]
-for i,c in zip(field, p):
-    i, c = int(i), int(c)
-    predicted_values[i-1] = c
-    row = int((i-1)/9)
-    col = int((i-1)%9)
-    print i, row, col, c
-    field_param += ' '+str(row)+str(col)+str(c)
-    rows[row].append(c)
-    cols[col].append(c)
-
-# accuracy
-section('Accuracy')
-# filter(lambda a: a 0= 0, predicted_values)
 solution = [
 3,0,0, 2,4,0, 0,6,0,
 0,4,0, 0,0,0, 0,5,3,
@@ -56,12 +32,54 @@ solution = [
 0,0,1, 9,2,0, 5,0,0,
 2,0,0, 3,0,0, 7,4,0,
 9,6,0, 5,0,0, 3,0,2]
-correct = [a for (a,b) in zip(solution,predicted_values) if a==b]
-failed  = [(i+1,a,b) for (i,a,b) in zip(range(9*9), solution,predicted_values) if a!=b]
-print 'Correct: ', len(correct), 'Failed: ', len(failed)
-print failed
-print 'Accuracy: ', len(correct)/(9.*9)
+def checkAccuracy(pred, printFailed=False):
[…]

### 2012-03-15 doc/icml2010stylefiles/paper.tex
- fixed/improved svm grid search
- added grid search result plotting

--- 2012-03-15 doc/icml2010stylefiles/paper.tex
@@
 \end{align*}
-Using a radial basis function (RBF) as kernel we performed a grid search over $C$ and $\gamma$ to find those parameters that provide the highest accuracy on the validation set.
-We used the open source library ``libsvm'' \cite{chang2011libsvm} wrapped by some octave scripts for setting up data sets and performing the grid search.
+Using a radial basis function (RBF) as kernel we performed a grid search over $C$ and $\gamma$ to find those parameters that provide the highest accuracy on the test set.
+We used the open source library ``libsvm'' \cite{chang2011libsvm} wrapped by some octave scripts for setting up data sets and performing the grid search with k-fold cross validation.
+Grid search with cross validation resulted in accuracies illustrated in \ref{fig:svm_grid_search}, thus we chose $C = ?$ and $\lambda = ?$.
+\begin{figure}[ht]
+	\centerline{\includegraphics[scale=0.20, trim=0 0 10 0]{gridSearchSVMPlot}}
+	\caption{}
+	\label{fig:svm_grid_search}
+\end{figure}
+
 \subsection{Logistic Regression}
+
+\color{red}performance on an example sudoku\color{black}
 

### 2012-03-15 gen-data.py

@@ -11,7 +11,7 @@ if(concrete):
     fonts = ['comicbd.ttf', 'arialbd.ttf'] #, 'arialbd.ttf', 'ariblk.ttf']
     size_mu = 16
     size_sig = 2
-    pos_sig = 3
+    pos_sig = 2
     color_sig = 10
     rotate_sig = 0
 else:

### 2012-03-15 octave/gridSearchSVMPlotResult.m

@@ -1,0 +1,56 @@
+% plots data which was generated by gridSearchSVM
+clear ; close all; clc
+if !exist('gridSearchSVMResult.mat')
+	printf("no data to plot, first run gridSearchSVM!");
+endif
+load('gridSearchSVMResult.mat');
+
+% Create New Figure
+figure;
+%hold on;
+%hold off;
+
+C = [2^-5,  2^-3,  2^-1,  2^1,  2^3,  2^5,  2^7,  2^9,  2^11, 2^13];
+g = [2^-15, 2^-13, 2^-11, 2^-9, 2^-7, 2^-5, 2^-3, 2^-1, 2^1,  2^3];
+for i=1:length(C)
+	C(i) = log2(C(i));
+	g(i) = log2(g(i));
+end
+% plot(C,g,'rx','MarkerSize', 5);
+
+% generating sample data
+acc = rand(1,length(C)*length(C));
+acc = reshape(acc,length(C),length(C));
+acc *= 100;
+for i=1:size(acc,1)
+	for j=1:size(acc,2)
+		if (i < 4 || i>6 || j<4 || j>6)
+			acc(i,j) = max([20, acc(i,j)-40]);
+		else
+			acc(i,j) = min([80, acc(i,j)+40]);
+		endif
+	end
+end
+
+% interpolating with spline to get a smooth shaped plot of the function
+[X,Y] = meshgrid(C,g);
+[XI,YI] = meshgrid(min(C)-1:0.25:max(C)+1, min(g)-1:0.25:max(g)+1);
+ZI = interp2(X,Y,acc,XI,YI,'spline');
+%mesh(C,g,acc), hold, mesh(XI,YI,ZI);
+mesh(XI,YI,ZI);
+
+% set plot metadata
+fontsize = 20;
+title("Grid Search with Cross-Validation Accuracy", "fontsize", fontsize);
+ylabel("log_2(\gamma)", "fontsize", fontsize);
+xlabel("log_2(C)", "fontsize", fontsize);
+zlabel("Accuracy", "fontsize", fontsize);
+axis([-6 14 -16 4 0 100]);
+set(gca, 'XTick', -5:2:13);
+set(gca, 'YTick', -15:2:3);
+set(gca, 'ZTick', 0:10:100);
+set(gca, 'fontsize', 20);
+grid on;
+
+% save
+print("gridSearchSVMPlot.png");

### 2012-03-15 octave/main.m

@@ -105,4 +105,5 @@ if (do_cross_validation)
 endif
 
 % save the trained models
+printf('saving model...');
 save("-mat-binary", models_file, "Theta1", "Theta2", "model", "Theta");

### 2012-03-15 octave/config.m
refactoring: common naming scheme {method}{Train|Predict}
config: lambda is now a vector -> 3 lambdas for nn, svm, lr

@@ -3,9 +3,9 @@
 % if you change any value here, you have to call config()
 % within octave to regenerate the config.mat file
 
-function config (lambda=0.01, maxIter=100, gamma=1)
+function config (lambda=[1, 30, 20], maxIter=100, gamma=1)
 	% Setup parameters
-	data_folder = '../data';  % folder that contains the generated images
+	data_folder = '../data2';  % folder that contains the generated images
     cache_file  = 'cache.mat';% name of the cache file
 	models_file = 'models.mat'; % contains trained models/parameters of nn/svm/lr
 	input_folder = '../input'; % folder that contains all images from the observed sudoku
@@ -16,12 +16,12 @@ function config (lambda=0.01, maxIter=100, gamma=1)
 	hidden_layer_size = 25;   % 25 hidden units
 	num_labels = 9;           % numbers 1-9
 	
-	set_size = 1125;		  % size of the k sets
+	number_of_sets = 4;		  % # of k sets
 	do_cross_validation = false; % if set to false we train only one time (no cross validation)
 	
-	use_nn = true;
-	use_svm = false;
-	use_lr = false;
+	use_nn  = lambda(1) != -1;
+	use_svm = lambda(2) != -1;
+	use_lr  = lambda(3) != -1;
 	
 	save("-mat-binary", "config.mat");
 endfunction

### 2012-03-15 octave/logRegPredict.m

@@ -1,4 +1,4 @@
-function p = predictLogReg(Theta, X)
+function p = logRegPredict(Theta, X)
 %PREDICT Predict whether the label is 0 or 1 using learned logistic 
 %regression parameters theta
 %   p = PREDICT(theta, X) computes the predictions for X using a 

### 2012-03-15 octave/logRegTrain.m

@@ -1,4 +1,4 @@
-function theta = logReg (X,y)
+function theta = logRegTrain(X, y, _lambda, _maxIter)
 	load('config.mat');
 	
 	% Some useful variables
@@ -20,14 +20,14 @@ function theta = logReg (X,y)
 	initial_theta = zeros(n+1, 1);
 
 	% Set regularization parameter lambda to 1 (you should vary this)
-	lambda = 1;
+	% lambda = 1;
 
 	% Set Options
-	options = optimset('GradObj', 'on', 'MaxIter', maxIter);	
+	options = optimset('GradObj', 'on', 'MaxIter', _maxIter);	
 
 	for i=1:9
 		printf("logistic regression for label %i: \n", i);
-		theta(:,i) = fmincg (@(t)(costFunctionReg(t, X, (y == i), lambda)), initial_theta, options);
+		theta(:,i) = fmincg (@(t)(costFunctionReg(t, X, (y == i), _lambda)), initial_theta, options);
 	end
 end
 

### 2012-03-15 octave/main.m

@@ -6,6 +6,7 @@ if !exist('config.mat')
 endif
 load('config.mat');
 lambda, maxIter, gamma
+fprintf('The following models will be added/overwritten:\n')
 use_nn, use_svm, use_lr
 if exist(models_file)
     load(models_file);
@@ -32,6 +33,7 @@ else
     %save(cache, "X", "y");
 endif
 X(:,size(X,2)+1) = y;
+n_samples = length(y)
 
 % mix row indexes
 all_indexed_mixed = randperm(size(X,1));
@@ -42,7 +44,7 @@ average_accuracy_log_reg = 0;
 cpu_time_nn = 0;
 cpu_time_svm = 0;
 cpu_time_lr = 0;
-number_of_sets = ceil(length(y)/set_size)
+set_size = ceil(length(y)/number_of_sets)
 
 if (do_cross_validation)
 	loops = number_of_sets-1;
@@ -65,11 +67,11 @@ for i = 0:loops
 		printf('\n=== neural network in iteration %i\n', i);
 		% neural network
 		t=cputime;
-		[Theta1, Theta2] = train(X_train,y_train);
+		[Theta1, Theta2] = nnTrain(X_train, y_train, lambda(1), maxIter);
 		cpu_time_nn += cputime-t;
 		%printf('Total cpu time for training: %f seconds\n', cputime-t);
 		% accuracy when predicting values of test set
-		pred = predictNN(Theta1, Theta2, X_val);
+		pred = nnPredict(Theta1, Theta2, X_val);
 		acc = mean(double(pred == y_val))*100;
 		fprintf('Test Set Accuracy (Neural Network): %f\n', acc);
 		average_accuracy += length(y_val)*acc;
@@ -79,7 +81,7 @@ for i = 0:loops
 		printf('\n=== svm in iteration %i\n', i);
 		% svm
 		t=cputime;
-		model = svmtrain(y_train, X_train, sprintf("-q -c %f -t 2 -g %f", lambda, gamma));
+		model = svmtrain(y_train, X_train, sprintf("-q -c %f -t 2 -g %f", lambda(2), gamma));
 		cpu_time_svm += cputime-t;
 		%printf('Total cpu time for training: %f seconds\n', cputime-t);
 		fprintf('Test Set Accuracy (SVM): ');
@@ -91,10 +93,10 @@ for i = 0:loops
 		printf('\n=== logistic regression in iteration %i\n', i);
 		% logistic regression
[…]

### 2012-03-15 octave/nnPredict.m

@@ -1,4 +1,4 @@
-function p = predictNN(Theta1, Theta2, X)
+function p = nnPredict(Theta1, Theta2, X)
 %PREDICT Predict the label of an input given a trained neural network
 %   p = PREDICT(Theta1, Theta2, X) outputs the predicted label of X given the
 %   trained weights of a neural network (Theta1, Theta2)

### 2012-03-15 octave/nnTrain.m

@@ -1,4 +1,4 @@
-function [Theta1, Theta2] = train(X, y)
+function [Theta1, Theta2] = nnTrain(X, y, _lambda, _maxIter)
 	load('config.mat');
 	
 	m = size(X, 1);
@@ -22,15 +22,15 @@ function [Theta1, Theta2] = train(X, y)
 	fprintf('Training Neural Network (Backpropagation)...\n');
 
 	% Iterations
-	options = optimset('MaxIter', maxIter);
+	options = optimset('MaxIter', _maxIter);
 
-	fprintf('Using lambda=%f for regularization\n', lambda)
+	fprintf('Using lambda=%f for regularization\n', _lambda)
 
 	% Create "short hand" for the cost function to be minimized
 	costFunction = @(p) nnCostFunction(p, ...
 									   input_layer_size, ...
 									   hidden_layer_size, ...
-									   num_labels, X, y, lambda);
+									   num_labels, X, y, _lambda);
 
 	% Now, costFunction is a function that takes in only one argument (the
 	% neural network parameters)

### 2012-03-15 gatherData.py
gatherData2: enter solution, multiple captures automatically classified
refactoring: naming, config
gitignore

@@ -1,0 +1,71 @@
+import os, sys, glob, shutil
+import Image, ImageDraw, ImageFont, ImageStat
+import numpy, pylab
+import matplotlib.pyplot as plt
+
+ext    = '.png'
+input  = 'input/*'+ext
+output = 'data2/'
+cache  = output+'cache.mat'
+cv_img = 'cv/sudoku.png'
+
+
+def show(img):
+    # imgplot = plt.imshow(img)
+    # plt.show()
+    img.show() # PIL
+
+def number_input(msg=''):
+    while True:
+        try:
+            c = int(raw_input('Enter number: ' if not len(msg) else msg))
+            if c < 0 or c > 9: raise ValueError
+            return c
+        except ValueError:
+            print 'Number has to be 0-9 (0 for empty fields)!'
+
+def getNewFilename(c):
+    dst = output+str(c)+'/'
+    files = [f for f in glob.glob(dst+'*'+ext)]
+    if len(files):
+        i = [int(os.path.basename(f).replace(ext, '')) for f in files]
+        i.sort()
+        i = i[-1]+1
+    else:
+        i = 0
+    return dst+str(i)+ext
+
+def classify():
+    print 'Please classify (0-9), 0 for empty field.'
+    plt.show()
+    for f in glob.glob(input):
+        img = Image.open(f)
+        show(img)
+        c = number_input()
+        ff = getNewFilename(c)
+        print 'move', f, ff
+        shutil.move(f, ff)
+
+def main():
+    if not os.path.exists(output):
+        [os.makedirs(output+str(i)) for i in range(0,10)]
+    print "This program will (in a loop) start cv.exe to capture an image (hit 'c'), split it and then ask you to classify the numbers."
+    print "Enter 0 for empty fields"
+    run = True
+    while run:
+        # delete existing capture so that the same samples aren't added twice
+        if os.path.exists(cv_img):
+            os.remove(cv_img)
+        else:
+            if raw_input("No input image found. Press 'x' to quit, anything else to continue and capture one.")=='x': break
+        # fresh image from capture
+        os.system('run-cv.bat')
+        # split image and save in input
+        os.system('python splitImage.py')
+        # delete cache before any new samples are added
+        if os.path.exists(cache): os.remove(cache)
+        # ask user to classify images until input is empty
+        classify()
+
+if __name__ == '__main__':
+    main()

### 2012-03-15 octave/config.m

@@ -6,7 +6,7 @@
 function config (lambda=[1, 30, 20], maxIter=100, gamma=1)
 	% Setup parameters
 	data_folder = '../data2';  % folder that contains the generated images
-    cache_file  = 'cache.mat';% name of the cache file
+    cache_file  = 'cache.mat'; % name of the cache file
 	models_file = 'models.mat'; % contains trained models/parameters of nn/svm/lr
 	input_folder = '../input'; % folder that contains all images from the observed sudoku
 	%lambda = 1;               % Regularization

### 2012-03-15 octave/predict.m

@@ -5,11 +5,11 @@ function [pNN, pSVM, pLR] = predict(X)
 	pLR = ones(m,1)*(-1);
 	load('config.mat');
 	load(models_file);
-	
+
 	
 	if (exist("Theta1") && exist("Theta2"))
 		% predict with NN
-		pNN = predictNN(Theta1, Theta2, X);
+		pNN = nnPredict(Theta1, Theta2, X);
 	endif
 	if (exist("model"))
 		% predict with SVM		
@@ -18,7 +18,7 @@ function [pNN, pSVM, pLR] = predict(X)
 		[pSVM,_,_] = svmpredict(ones(m,1),X,model);
 	endif
 	if (exist("Theta"))
-		pLR = predictLogReg(Theta,X);
+		pLR = logRegPredict(Theta,X);
 	endif
     
     % test: show images and predicted class

### 2012-03-15 octave/processInput.m

@@ -7,6 +7,7 @@ X = []; % matrix to store the image vectors
 field = []; % vector that stores the corresponding sudoku field of a row in X matrix
 
 files = glob(sprintf('%s%s', path, '*.png'));
+
 for x=1:length(files)
 	if(!isdir(files{x}))
 		rgbImg = imread(files{x});

### 2012-03-15 splitImage.py

@@ -3,7 +3,7 @@ import random, os, shutil
 
 input = 'cv/sudoku.png'
 out = 'input/'
-tmp = out+'tmp/'
+tmp = out+'all/'
 targetsize = 20
 m = 0
 threshold = 250 # brighter fields are empty

### 2012-03-15 mahotas.py
cleanup

@@ -1,11 +1,0 @@
-import numpy as np
-from scipy import ndimage
-import mahotas
-import pylab
-
-img = mahotas.imread('1.png')
-T_otsu = mahotas.thresholding.otsu(img)
-seeds,_ = ndimage.label(img > T_otsu)
-labeled = mahotas.cwatershed(img.max() - img, seeds)
-
-pylab.imshow(labeled)

### 2012-03-15 nn.py
batch-files
gitignore

@@ -24,7 +24,7 @@ y = mat['y']
 n_in = int(config['input_layer_size'][0][0])
 n_hidden = int(config['hidden_layer_size'][0][0])
 n_classes = int(config['num_labels'][0][0])
-weightdecay = config['lambda'][0][0]
+weightdecay = config['lambda'][0][0]-1
 maxIter = int(config['maxIter'][0][0])
 
 from pybrain.tools.customxml import NetworkWriter

### 2012-03-15 octave/config.m
history of models:
d1 = generated, d2 = gathered
a_b_c -> lambdas for nn, svm, lr

@@ -3,9 +3,9 @@
 % if you change any value here, you have to call config()
 % within octave to regenerate the config.mat file
 
-function config (lambda=[1, 30, 20], maxIter=100, gamma=1)
+function config (lambda=[1, 50, 20], maxIter=100, gamma=1)
 	% Setup parameters
-	data_folder = '../data2';  % folder that contains the generated images
+	data_folder = '../data';  % folder that contains the generated images
     cache_file  = 'cache.mat'; % name of the cache file
 	models_file = 'models.mat'; % contains trained models/parameters of nn/svm/lr
 	input_folder = '../input'; % folder that contains all images from the observed sudoku

### 2012-03-15 octave/config.m
models: data3: data and data2 combined

@@ -5,7 +5,7 @@
 
 function config (lambda=[1, 50, 20], maxIter=100, gamma=1)
 	% Setup parameters
-	data_folder = '../data';  % folder that contains the generated images
+	data_folder = '../data2';  % folder that contains the generated images
     cache_file  = 'cache.mat'; % name of the cache file
 	models_file = 'models.mat'; % contains trained models/parameters of nn/svm/lr
 	input_folder = '../input'; % folder that contains all images from the observed sudoku

### 2012-03-15 doc/icml2010stylefiles/paper.tex
fixed problem with hyperref
- finished gridSearchSVM to work with the real data
- worked on doc
- moved resource files from doc one folder up
- removed ex4.pdf

--- 2012-03-15 doc/icml2010stylefiles/paper.tex
@@
 % \usepackage[nohyperref]{icml2010} above.
-\usepackage{hyperref}
+%\usepackage{hyperref}
 
@@
 % note in the first column to ``Appearing in''
-\usepackage[accepted]{icml2010}
+\usepackage[accepted, nohyperref]{icml2010}
 
@@
 Digitalizing print media, automated sorting of letters, or recognition of signs in robotics or driver assistance systems require well performing algorithms for character recognition.
-In the following we analyze the performance of different techniques for number recognition, namely Neural Networks, Support Vector Machines and Logistic Regression, by using them in an automated Sudoku solver.
+In the following we analyze the performance of different techniques for number recognition, namely Neural Networks (NN), Support Vector Machines (SVM) and Logistic Regression (LR), by using them in an automated Sudoku solver.
 
 \section{Overview}
+Figure \ref{fig:overview} shows the basic data flow and the participating data manipulators.
+First of all we generate data by sampling position, rotation, size and color of a number and place it based on these values on a white 20x20 pixel image.
+Examples of 9 such generated images are shown in Figure \ref{fig:generated_numbers}.
+These images are preprocessed by an octave script and written to a matrix.
+The corresponding labels are defined by the name of the folder the images are stored in.
 % trim=l b r t
 \begin{figure}[ht]
-	\centerline{\includegraphics[width=\columnwidth, trim=25 30 25 30]{overview}}
+	\centerline{\includegraphics[width=\columnwidth, trim=25 30 25 30]{../overview}}
 	\caption{Overview}
@@
 \end{figure}
+The $X^{n\text{x}400}$ and the target vector $y$ are the inputs for training NN, SVM and LR, as described in Section \ref{training}.
+Trained models/parameters are saved in ``.mat'' files that can be loaded by the number prediction.
+The other input for number prediction is an input matrix of the same form as $X$ with a row entry for each Sudoku field that was detected to contain a number.
 
@@
 \begin{figure}[ht]
-	\centerline{\includegraphics[scale=0.25]{examples}}
+	\centerline{\includegraphics[scale=0.25]{../examples}}
 	\caption{Examples of generated numbers}
@@
 \section{Training}
+\label{training}
 \subsection{Neural Network}
+Applying a NN seems promising as it was already successfully applied to the quite similar problem of detecting handwritten digits.
+\color{red}ref!\color{black} For the NN we choose a simple architecture with 25 hidden sigmoidal units (+ bias unit), as shown in figure \ref{fig:nn}.
+The size of the input layer is 400 (given by the 20 x 20 pixel images) and the size of the output layer is defined by the number of classes, in this case the numbers 1 to 9.
+This simple architecture turned out to perform pretty well compared to more complex ones with a higher number of hidden layers.
+Training is performed with error backpropagation, thus minimizing the error function
+\begin{align*}
+& E(w) = \sum_{n=1}^N E_n(w)\\
+\text{with}~~E&_n(w) = \frac{1}{2}\sum_k (y_{nk}-t_{nk})\\
+\text{and}~~~~~&~~~y_k = \sum_i w_{ki}x_{i}\\
+\end{align*}
+subject to $w$.
 \begin{figure}[ht]
-	\centerline{\includegraphics[width=\columnwidth, trim=0 30 0 30]{nn}}
+	\centerline{\includegraphics[width=\columnwidth, trim=0 30 0 30]{../nn}}
 	\caption{Neural Network with one hidden layer}
@@
 \end{figure}
+For NNs we test two implementations.
+On the one hand we use a partly self-developed octave code for two-layer NNs which was written as part of the stanford machine learning course\footnote{http://www.ml-class.org}, on the other hand we include PyBrain \cite{schaul2010pybrain} for comparison and easier evaluation of more sophisticated network architectures.
+To find a good value for the regularization parameter $\lambda$ we perform a grid search with k-fold cross-validation which gives us a good performance when choosing $\lambda = 1$.
+
 \subsection{Support Vector Machines}
-The next technique we applied to the discussed problem are Support Vector Machines \cite{Bishop200608} which try to solve the following optimization problem:
+The next technique we apply to the discussed problem are Support Vector Machines \cite{Bishop200608} which try to solve the following optimization problem:
 \begin{align*}
@@
 \end{align*}
-Using a radial basis function (RBF) as kernel we performed a grid search over $C$ and $\gamma$ to find those parameters that provide the highest accuracy on the validation set.
-We used the open source library ``libsvm'' \cite{chang2011libsvm} wrapped by some octave scripts for setting up data sets and performing the grid search.
+Using a radial basis function (RBF) as kernel we perform a grid search over $C$ and $\gamma$ to find those parameters that provide highest accuracy on the test set.
+We use the open source library ``libsvm'' \cite{chang2011libsvm} wrapped by some octave scripts for setting up data sets and performing grid search with k-fold cross-validation.
+Grid search with cross validation resulted in accuracies illustrated in Figure \ref{fig:svm_grid_search}.
+From the best parameter pairs we choose the one that performs best on the validation set, which gives us \color{red}$C = ?$ and $\lambda = ?$.\color{black}
+\begin{figure}[ht]
+	\centerline{\includegraphics[scale=0.20, trim=0 0 10 0]{../gridSearchSVMPlot}}
+	\caption{}
+	\label{fig:svm_grid_search}
+\end{figure}
+
 \subsection{Logistic Regression}
+Finally we apply Logistic Regression to the datasets. Again we use a (vectorized) octave implementation for learning and prediction. Therefore we minimize the (regularized) error function
+\begin{align*}
+E(w) &= \frac{1}{n}(-y^T ln(\Phi w) + (1-y^T)ln(1-\Phi w))\\
+&~~~+\frac{\lambda}{2n}w^Tw
+\end{align*}
+with respect to $w$.
+Minimization is performed by an algorithm\footnote{http://sprinkler.googlecode.com/svn/trunk/\\regression/fmincg.m} that was provided during the stanford machine learning course.
+Grid search and k-fold cross-validation recommend a value for the regularization parameter $\lambda$.
 
 \section{Image Recognition and Preprocessing}
-
-\section{Number Prediction}
+Image recognition is written in C by extensive use of the open source library OpenCV\footnote{http://opencv.willowgarage.com/wiki/}.
+Images captured by the webcam are continously analyzed to find a Sudoku square.
+Images are transformed to greyscale, then an adaptive threshold algorithm is used to preprocess the image for contour detection by OpenCV.
+OpenCV returns a list of contours which are further checked for certain criteria like having four corners or being of certain size.
+Since the contours are not detected precise enough lines are fitted through six checkpoints on each edge in order to achieve subpixel accuracy.
+The content enclosed by the corners is perspectively warped to fit a predefined square size.
+Finally, the thresholded Sudoku field is saved to disk.
+A python script is then used to split the whole field into 9x9 subfields with each subfield being flood filled around the corners for noise reduction.
+The format of the subfields now equals to the one provided by the generated images which allows us to apply the trained NN/SVM/LR models and, finally, apply a Sudoku solver.
 
 \section{Evaluation}
+For evaluation of the different techniques we define the following datasets:
+\begin{itemize}
+\item d1: 4500 generated images
+\item d2: $\sim$ 1000 gathered images
+\item d3: D1 $\cup$ D2
+\item dVal: images of a captured Sudoku
+\end{itemize}
+
+\color{red}
+\begin{table}[t]
+\caption{Classification accuracies for Neural Network, SVM and Logistic Regression on different data sets.}
+\label{sample-table}
+\vskip 0.15in
+\begin{center}
+\begin{small}
+\begin{sc}
+\begin{tabular}{lcccc}
+\hline
+\abovespace\belowspace
+Data set & NN & SVM & LR & Winner \\
+\hline
+\abovespace
+D1		& 95.9$\pm$ 0.2& 96.7$\pm$ 0.2& 96.7$\pm$ 0.2& ?\\
+D2		& 83.3$\pm$ 0.6& 80.0$\pm$ 0.6& ? & ?\\
+D3		& 61.9$\pm$ 1.4& 83.8$\pm$ 0.7& ? & ? \\
+\belowspace
+DVal	& 74.8$\pm$ 0.5& 78.3$\pm$ 0.6& ? & ?        \\
+\hline
+\end{tabular}
+\end{sc}
+\end{small}
+\end{center}
+\vskip -0.1in
+\end{table}
+\color{black}
 
 \section{Future Work}
-\color{red} include information if a sudoku is unsolvable \color{black}
+\color{red}
+\begin{itemize}
+\item include information if a sudoku is unsolvable
+\item online learning
+\end{itemize}
+\color{black}
 

### 2012-03-15 octave/gridSearchSVM.m

@@ -4,9 +4,6 @@
 C = [2^-5,  2^-3,  2^-1,  2^1,  2^3,  2^5,  2^7,  2^9,  2^11, 2^13];
 g = [2^-15, 2^-13, 2^-11, 2^-9, 2^-7, 2^-5, 2^-3, 2^-1, 2^1,  2^3];
 
-%C = [2^5]
-%g = [2^-5]
-
 if !exist('config.mat')
 	config();
 endif

### 2012-03-15 octave/gridSearchSVMPlotResult.m

@@ -10,30 +10,35 @@ figure;
 %hold on;
 %hold off;
 
-C = [2^-5,  2^-3,  2^-1,  2^1,  2^3,  2^5,  2^7,  2^9,  2^11, 2^13];
-g = [2^-15, 2^-13, 2^-11, 2^-9, 2^-7, 2^-5, 2^-3, 2^-1, 2^1,  2^3];
+%C = [2^-5,  2^-3,  2^-1,  2^1,  2^3,  2^5,  2^7,  2^9,  2^11, 2^13];
+%g = [2^-15, 2^-13, 2^-11, 2^-9, 2^-7, 2^-5, 2^-3, 2^-1, 2^1,  2^3];
 for i=1:length(C)
 	C(i) = log2(C(i));
 	g(i) = log2(g(i));
 end
 % plot(C,g,'rx','MarkerSize', 5);
 
+% values in acc were added by iterating through C and iterating through g in an inner loop
+% thus, the first length(C) values are for C(1) and g(1)-g(length(g))
+[X,Y] = meshgrid(C,g);	% X has same C values in one column, Y has same g values in row
+% so we can just reshape the acc vector because it will iteratively fill the columns
+acc = reshape(acc, length(C), length(g));
+
 % generating sample data
-acc = rand(1,length(C)*length(C));
-acc = reshape(acc,length(C),length(C));
-acc *= 100;
-for i=1:size(acc,1)
-	for j=1:size(acc,2)
-		if (i < 4 || i>6 || j<4 || j>6)
-			acc(i,j) = max([20, acc(i,j)-40]);
-		else
-			acc(i,j) = min([80, acc(i,j)+40]);
-		endif
-	end
-end
+%acc = rand(1,length(C)*length(C));
+%acc = reshape(acc,length(C),length(C));
+%acc *= 100;
+%for i=1:size(acc,1)
+%	for j=1:size(acc,2)
+%		if (i < 4 || i>6 || j<4 || j>6)
+%			acc(i,j) = max([20, acc(i,j)-40]);
+%		else
+%			acc(i,j) = min([80, acc(i,j)+40]);
+%		endif
+%	end
+%end
 
 % interpolating with spline to get a smooth shaped plot of the function
-[X,Y] = meshgrid(C,g);
 [XI,YI] = meshgrid(min(C)-1:0.25:max(C)+1, min(g)-1:0.25:max(g)+1);
[…]

### 2012-03-15 octave/config.m
all models 100%

@@ -3,7 +3,7 @@
 % if you change any value here, you have to call config()
 % within octave to regenerate the config.mat file
 
-function config (lambda=[1, 50, 20], maxIter=100, gamma=1)
+function config (lambda=[1, 20, 20], maxIter=100, gamma=.1)
 	% Setup parameters
 	data_folder = '../data2';  % folder that contains the generated images
     cache_file  = 'cache.mat'; % name of the cache file

### 2012-03-15 doc/icml2010stylefiles/paper.tex
worked on doc

@@
 % note in the first column to ``Appearing in''
-\usepackage[accepted, nohyperref]{icml2010}
+%\usepackage[accepted, nohyperref]{icml2010}
+\usepackage[nohyperref]{icml2010}
 
@@
 First of all we generate data by sampling position, rotation, size and color of a number and place it based on these values on a white 20x20 pixel image.
-Examples of 9 such generated images are shown in Figure \ref{fig:generated_numbers}.
-These images are preprocessed by an octave script and written to a matrix.
-The corresponding labels are defined by the name of the folder the images are stored in.
+Examples of 9 such generated images are shown in Figure \ref{fig:generated_numbers}.The corresponding labels are defined by the name of the folder the images are stored in.
+\begin{figure}[ht]
+	\centerline{\includegraphics[scale=0.25]{../examples}}
+	\caption{Examples of generated numbers}
+	\label{fig:generated_numbers}
+\end{figure}
+These images are preprocessed (applying greyscale) by an octave script and written to an input matrix and a target vector.
+All pixels of one image are encoded in one row of the input matrix and the corresponding label is stored in the row of the target vector.
+The data is dumped to a ``cache.mat'' file which allows us to apply many different tools and programming languages to the dataset.
+In our case we use it in python and octave.
 % trim=l b r t
@@
 \end{figure}
-The $X^{n\text{x}400}$ and the target vector $y$ are the inputs for training NN, SVM and LR, as described in Section \ref{training}.
-Trained models/parameters are saved in ``.mat'' files that can be loaded by the number prediction.
+The input matrix $X^{n\text{x}400}$ and the target vector $y$ are the inputs for training NN, SVM and LR, as described in Section \ref{training}.
+After training, the learned models/parameters are saved in ``.mat'' files that can be loaded by the number prediction.
 The other input for number prediction is an input matrix of the same form as $X$ with a row entry for each Sudoku field that was detected to contain a number.
-
-\section{Generation of Training Data}
-\subsection{Image Generation}
-\begin{figure}[ht]
-	\centerline{\includegraphics[scale=0.25]{../examples}}
-	\caption{Examples of generated numbers}
-	\label{fig:generated_numbers}
-\end{figure}
-\subsection{Image Processing}
+This implies that we do not train our models to detect empty fields.
+We could reliably detect empty fields by performing some preprocessing steps on the images (filling borders, adjusting thresholds etc.) thus applying a machine learning technique does not seem appropriate for this task.
 
@@
 Grid search with cross validation resulted in accuracies illustrated in Figure \ref{fig:svm_grid_search}.
-From the best parameter pairs we choose the one that performs best on the validation set, which gives us \color{red}$C = ?$ and $\lambda = ?$.\color{black}
+From the best parameter pairs we choose the one that performs best on the validation set, which gives us $C = 20$ and $\gamma = 0.1$.
 \begin{figure}[ht]
@@
 \section{Evaluation}
-For evaluation of the different techniques we define the following datasets:
+To evaluate the performance of different approaches we define the following data sets:
 \begin{itemize}
@@
 
-\section{Electronic Submission}
-\label{submission}
-
-As in the past few years, ICML 2010 will rely exclusively on
-electronic formats for submission and review. We assume that all
-authors will have access to standard software for word processing,
-electronic mail, and web file transfer. Authors who do not have such
-access should send email with their concerns to
-\texttt{icml10.programchairs@gmail.com}.
-
-\subsection{Templates for Papers}
-
-Electronic templates for producing papers for submission are available
-for \LaTeX\/ and Microsoft Word. Templates are accessible on the World
-Wide Web at:\\
-\textbf{\texttt{http://www.icml2010.org/}}
-
-\noindent
-Send questions about these electronic templates to
-\texttt{icml10.programchairs@gmail.com}.
-
-If you are using the templates, the formatting instructions below
-will already be enforced. The content rules you must follow are:
-\begin{itemize}
-\item The maximum paper length is 8 pages.
-\item Do not include author information or acknowledgments in your
-  initial submission. 
-\item Place figure captions {\em under} the figure (and omit titles from
-  inside the graphic file itself).  Place table captions {\em over}
-  the table.
-\item References must include page numbers whenever possible and be as
-  complete as possible.  Place multiple citations in chronological order.  
-\end{itemize}
-Please see below for details on each of these items.
-
-\subsection{Submitting Papers}
-
-Submission to ICML 2010 will be entirely electronic, via a web site
-(not email).  The URL and information about the submission process
-are available on the conference web site at
-
-\textbf{\texttt{http://www.icml2010.org/sub.html}}
-
-{\bf Paper Deadline:} The deadline for paper submission to ICML 2010
-is Monday, February 1, 2010, at 11:59 p.m.~Samoa time.  If your full
-submission does not reach us by this date, it will not be considered
-for publication. There is no separate abstract submission this year.
-
-{\bf Anonymous Submission:} To facilitate blind review, no identifying
-author information should appear on the title page or in the paper
-itself.  Section~\ref{author info} will explain the details of how to
-format this.
-
-{\bf Simultaneous Submission:} ICML will not accept any paper which,
-at the time of submission, is under review for another conference or
-has already been published. This policy also applies to papers that
-overlap substantially in technical content with conference papers
-under review or previously published. ICML submissions must not be
-submitted to other conferences during ICML's review period. Authors
-may submit to ICML substantially different versions of journal papers
-that are currently under review by the journal, but not yet accepted
-at the time of submission. Informal publications, such as technical
-reports or papers in workshop proceedings which do not appear in
-print, do not fall under these restrictions.
-
-\medskip
-
-To ensure our ability to print submissions, authors must provide their
-manuscripts in \textbf{PDF} format.  Furthermore, please make sure
-that files contain only Type-1 fonts (e.g.,~using the program {\tt
-  pdffonts} in linux or using File/DocumentProperties/Fonts in
-Acrobat).  Other fonts (like Type-3) might come from graphics files
-imported into the document.
-
-Authors using \textbf{Word} must convert their document to PDF.  Most
-of the latest versions of Word have the facility to do this
-automatically.  Submissions will not be accepted in Word format or any
-format other than PDF. Really. We're not joking. Don't send Word.
-
-Those who use \textbf{\LaTeX} to format their accepted papers need to
-pay close attention to the typefaces used.  Specifically, when
-producing the PDF by first converting the dvi output of \LaTeX\ to Postscript
-the default behavior is to use non-scalable Type-3 PostScript bitmap
-fonts to represent the standard \LaTeX\ fonts. The resulting document
-is difficult to read in electronic form; the type appears fuzzy. To
-avoid this problem, dvips must be instructed to use an alternative
-font map.  This can be achieved with
-something like the following commands:\\[0.5em]
-{\bf dvips -Ppdf -tletter -G0 -o paper.ps paper.dvi}\\
-{\bf ps2pdf paper.ps}\\[0.5em]
-Note that it is a zero following the ``-G''.  This tells dvips to use
-the config.pdf file (and this file refers to a better font mapping).
-
-Another alternative is to use the \textbf{pdflatex} program instead of
-straight \LaTeX. This program avoids the Type-3 font problem, however
-you must ensure that all of the fonts are embedded (use {\tt
-pdffonts}). If they are not, you need to configure pdflatex to use a
-font map file that specifies that the fonts be embedded. Also you
-should ensure that images are not downsampled or otherwise compressed
-in a lossy way.
-
-Note that the new 2010 style files use the {\tt hyperref} package to
-make clickable links in documents.  If this causes problems for you,
-add {\tt nohyperref} as one of the options to the {\tt icml2010}
-usepackage statement.
-
-\subsection{Reacting to Reviews}
-We will continue the ICML tradition in which the authors are given the
-option of providing a short reaction to the initial reviews. These
-reactions will be taken into account in the discussion among the
-reviewers and area chairs.
-
-\subsection{Submitting Final Camera-Ready Copy}
-
-The final versions of papers accepted for publication should follow the
-same format and naming convention as initial submissions, except of
-course that the normal author information (names and affiliations)
-should be given.  See Section~\ref{final author} for details of how to
-format this.
-
-The footnote, ``Preliminary work.  Under review by the International
-Conference on Machine Learning (ICML).  Do not distribute.'' must be
-modified to ``Appearing in \textit{Proceedings of the
-$\mathit{27}^{th}$ International Conference on Machine Learning},
-Haifa, Israel, 2010.  Copyright 2010 by the author(s)/owner(s).''
-
-For those using the \textbf{\LaTeX} style file, simply change
-$\mathtt{\backslash usepackage\{icml2010\}}$ to
-
-\verb|\usepackage[accepted]{icml2010}|
-
-\noindent
-Authors using \textbf{Word} must edit the
-footnote on the first page of the document themselves.
-
-Camera-ready copies should have the title of the paper as running head
-on each page except the first one.  The running title consists of a
-single line centered above a horizontal rule which is $1$ point thick.
-The running head should be centered, bold and in $9$ point type.  The
-rule should be $10$ points above the main text.  For those using the
-\textbf{\LaTeX} style file, the original title is automatically set as running
-head using the {\tt fancyhdr} package which is included in the ICML
-2010 style file package.  In case that the original title exceeds the
-size restrictions, a shorter form can be supplied by using
-
-\verb|\icmltitlerunning{...}|
-
-just before $\mathtt{\backslash begin\{document\}}$.
-Authors using \textbf{Word} must edit the header of the document themselves.
-
-\section{Format of the Paper}
-
-All submissions must follow the same format to ensure the printer can
-reproduce them without problems and to let readers more easily find
-the information that they desire.
-
-\subsection{Length and Dimensions}
-
-Papers must not exceed eight (8) pages, including all figures, tables,
-references, and appendices. Any submission that exceeds this page
-limit or that diverges significantly from the format specified herein
-will be rejected without review.
-
-The text of the paper should be formatted in two columns, with an
-overall width of 6.75 inches, height of 9.0 inches, and 0.25 inches
-between the columns. The left margin should be 0.75 inches and the top
-margin 1.0 inch (2.54~cm). The right and bottom margins will depend on
-whether you print on US letter or A4 paper, but all final versions
-must be produced for US letter size.
-
-The paper body should be set in 10~point type with a vertical spacing
-of 11~points. Please use Times Roman typeface throughout the text.
-
-\subsection{Title}
-
-The paper title should be set in 14~point bold type and centered
-between two horizontal rules that are 1~point thick, with 1.0~inch
-between the top rule and the top edge of the page. Capitalize the
-first letter of content words and put the rest of the title in lower
-case.
-
-\subsection{Author Information for Submission}
-\label{author info}
-
-To facilitate blind review, author information must not appear.  If
-you are using \LaTeX\/ and the \texttt{icml2010.sty} file, you may use
-\verb+\icmlauthor{...}+ to specify authors.  The author information
-will simply not be printed until {\tt accepted} is an argument to the
-style file. Submissions that include the author information will not
-be reviewed.
-
-\subsubsection{Self-Citations}
-
-If your are citing published papers for which you are an author, refer
-to yourself in the third person. In particular, do not use phrases
-that reveal your identity (e.g., ``in previous work \cite{langley00}, we 
-have shown \ldots'').
-
-Do not anonymize citations in the reference section by removing or
-blacking out author names. The only exception are manuscripts that are
-not yet published (e.g. under submission). If you choose to refer to
-such unpublished manuscripts \cite{anonymous}, anonymized copies have to be submitted
-as Supplementary Material via CMT. However, keep in mind that an ICML
-paper should be self contained and should contain sufficient detail
-for the reviewers to evaluate the work. In particular, reviewers are
-not required to look a the Supplementary Material when writing their
-review.
-
-\subsubsection{Camera-Ready Author Information}
-\label{final author}
-
-If a paper is accepted, a final camera-ready copy must be prepared.
-%
-For camera-ready papers, author information should start 0.3~inches
-below the bottom rule surrounding the title. The authors' names should
-appear in 10~point bold type, electronic mail addresses in 10~point
-small capitals, and physical addresses in ordinary 10~point type.
-Each author's name should be flush left, whereas the email address
-should be flush right on the same line. The author's physical address
-should appear flush left on the ensuing line, on a single line if
-possible. If successive authors have the same affiliation, then give
-their physical address only once.
-
-A sample file (in PDF) with author names is included in the ICML2010 style file package.
-
-\subsection{Abstract}
-
-The paper abstract should begin in the left column, 0.4~inches below
-the final address. The heading `Abstract' should be centered, bold,
-and in 11~point type. The abstract body should use 10~point type, with
-a vertical spacing of 11~points, and should be indented 0.25~inches
-more than normal on left-hand and right-hand margins. Insert
-0.4~inches of blank space after the body. Keep your abstract brief,
-limiting it to one paragraph and no more than six or seven sentences.
-
-\subsection{Partitioning the Text}
-
-You should organize your paper into sections and paragraphs to help
-readers place a structure on the material and understand its
-contributions.
-
-\subsubsection{Sections and Subsections}
-
-Section headings should be numbered, flush left, and set in 11~pt bold
-type with the content words capitalized. Leave 0.25~inches of space
-before the heading and 0.15~inches after the heading.
-
-Similarly, subsection headings should be numbered, flush left, and set
-in 10~pt bold type with the content words capitalized. Leave
-0.2~inches of space before the heading and 0.13~inches afterward.
-
-Finally, subsubsection headings should be numbered, flush left, and
-set in 10~pt small caps with the content words capitalized. Leave
-0.18~inches of space before the heading and 0.1~inches after the
-heading.
-
-Please use no more than three levels of headings.
-
-\subsubsection{Paragraphs and Footnotes}
-
-Within each section or subsection, you should further partition the
-paper into paragraphs. Do not indent the first line of a given
-paragraph, but insert a blank line between succeeding ones.
-
-You can use footnotes\footnote{For the sake of readability, footnotes
-should be complete sentences.} to provide readers with additional
-information about a topic without interrupting the flow of the paper.
-Indicate footnotes with a number in the text where the point is most
-relevant. Place the footnote in 9~point type at the bottom of the
-column in which it appears. Precede the first footnote in a column
-with a horizontal rule of 0.8~inches.\footnote{Multiple footnotes can
-appear in each column, in the same order as they appear in the text,
-but spread them across columns and pages if possible.}
-
-\begin{figure}[ht]
-\vskip 0.2in
-\begin{center}
-\centerline{\includegraphics[width=\columnwidth]{icml_numpapers}}
-\caption{Historical locations and number of accepted papers for International
-  Machine Learning Conferences (ICML 1993 -- ICML 2008) and
-  International Workshops on Machine Learning (ML 1988 -- ML
-  1992). At the time this figure was produced, the number of
-  accepted papers for ICML 2008 was unknown and instead estimated.}
-\label{icml-historical}
-\end{center}
-\vskip -0.2in
-\end{figure}
-
-\subsection{Figures}
-
-You may want to include figures in the paper to help readers visualize
-your approach and your results. Such artwork should be centered,
-legible, and separated from the text. Lines should be dark and at
-least 0.5~points thick for purposes of reproduction, and text should
-not appear on a gray background.
-
-Label all distinct components of each figure. If the figure takes the
-form of a graph, then give a name for each axis and include a legend
-that briefly describes each curve. Do not include a title inside the
-figure; instead, the caption should serve this function.
-
-Number figures sequentially, placing the figure number and caption
-{\it after\/} the graphics, with at least 0.1~inches of space before
-the caption and 0.1~inches after it, as in
-Figure~\ref{icml-historical}.  The figure caption should be set in
-9~point type and centered unless it runs two or more lines, in which
-case it should be flush left.  You may float figures to the top or
-bottom of a column, and you may set wide figures across both columns
-(use the environment {\tt figure*} in \LaTeX), but always place
-two-column figures at the top or bottom of the page.
-
-\subsection{Algorithms}
-
-If you are using \LaTeX, please use the ``algorithm'' and ``algorithmic'' 
-environments to format pseudocode. These require 
-the corresponding stylefiles, algorithm.sty and 
-algorithmic.sty, which are supplied with this package. 
-Algorithm~\ref{alg:example} shows an example.
-
-\begin{algorithm}[tb]
-   \caption{Bubble Sort}
-   \label{alg:example}
-\begin{algorithmic}
-   \STATE {\bfseries Input:} data $x_i$, size $m$
-   \REPEAT
-   \STATE Initialize $noChange = true$.
-   \FOR{$i=1$ {\bfseries to} $m-1$}
-   \IF{$x_i > x_{i+1}$} 
-   \STATE Swap $x_i$ and $x_{i+1}$
-   \STATE $noChange = false$
-   \ENDIF
-   \ENDFOR
-   \UNTIL{$noChange$ is $true$}
-\end{algorithmic}
-\end{algorithm}
-
-\subsection{Tables}
-
-You may also want to include tables that summarize material. Like 
-figures, these should be centered, legible, and numbered consecutively. 
-However, place the title {\it above\/} the table with at least 
-0.1~inches of space before the title and the same after it, as in 
-Table~\ref{sample-table}. The table title should be set in 9~point 
-type and centered unless it runs two or more lines, in which case it
-should be flush left.
-
-% Note use of \abovespace and \belowspace to get reasonable spacing 
-% above and below tabular lines.
-
-\begin{table}[t]
-\caption{Classification accuracies for naive Bayes and flexible 
-Bayes on various data sets.}
-\label{sample-table}
-\vskip 0.15in
-\begin{center}
-\begin{small}
-\begin{sc}
-\begin{tabular}{lcccr}
-\hline
-\abovespace\belowspace
-Data set & Naive & Flexible & Better? \\
-\hline
-\abovespace
-Breast    & 95.9$\pm$ 0.2& 96.7$\pm$ 0.2& $\surd$ \\
-Cleveland & 83.3$\pm$ 0.6& 80.0$\pm$ 0.6& $\times$\\
-Glass2    & 61.9$\pm$ 1.4& 83.8$\pm$ 0.7& $\surd$ \\
-Credit    & 74.8$\pm$ 0.5& 78.3$\pm$ 0.6&         \\
-Horse     & 73.3$\pm$ 0.9& 69.7$\pm$ 1.0& $\times$\\
-Meta      & 67.1$\pm$ 0.6& 76.5$\pm$ 0.5& $\surd$ \\
-Pima      & 75.1$\pm$ 0.6& 73.9$\pm$ 0.5&         \\
-\belowspace
-Vehicle   & 44.9$\pm$ 0.6& 61.5$\pm$ 0.4& $\surd$ \\
-\hline
-\end{tabular}
-\end{sc}
-\end{small}
-\end{center}
-\vskip -0.1in
-\end{table}
-
-Tables contain textual material that can be typeset, as contrasted 
-with figures, which contain graphical material that must be drawn. 
-Specify the contents of each row and column in the table's topmost
-row. Again, you may float tables to a column's top or bottom, and set
-wide tables across both columns, but place two-column tables at the
-top or bottom of the page.
-
-\subsection{Citations and References}
-
-Please use APA reference format regardless of your formatter
-or word processor. If you rely on the \LaTeX\/ bibliographic 
-facility, use {\tt natbib.sty} and {\tt icml2010.bst} 
-included in the style-file package to obtain this format.
-
-Citations within the text should include the authors' last names and
-year. If the authors' names are included in the sentence, place only
-the year in parentheses, for example when referencing Arthur Samuel's
-pioneering work \yrcite{Samuel59}. Otherwise place the entire
-reference in parentheses with the authors and year separated by a
-comma \cite{Samuel59}. List multiple references separated by
-semicolons \cite{kearns89,Samuel59,mitchell80}. Use the `et~al.'
-construct only for citations with three or more authors or after
-listing all authors to a publication in an earlier reference \cite{MachineLearningI}.
-
-Authors should cite their own work in the third person
-in the initial version of their paper submitted for blind review.
-Please refer to Section~\ref{author info} for detailed instructions on how to
-cite your own papers.
-
-Use an unnumbered first-level section heading for the references, and 
-use a hanging indent style, with the first line of the reference flush
-against the left margin and subsequent lines indented by 10 points. 
-The references at the end of this document give examples for journal
-articles \cite{Samuel59}, conference publications \cite{langley00}, book chapters \cite{Newell81}, books \cite{DudaHart2nd}, edited volumes \cite{MachineLearningI}, 
-technical reports \cite{mitchell80}, and dissertations \cite{kearns89}.
-
-Alphabetize references by the surnames of the first authors, with
-single author entries preceding multiple author entries. Order
-references for the same authors by year of publication, with the
-earliest first. Make sure that each reference includes all relevant
-information (e.g., page numbers).
-
-\subsection{Software and Data}
-
-We strongly encourage the publication of software and data with the
-camera-ready version of the paper whenever appropriate.  This can be
-done by including a URL in the camera-ready copy.  However, do not
-include URLs that reveal your institution or identity in your
-submission for review.  Instead, provide an anonymous URL or upload
-the material as ``Supplementary Material'' into the CMT reviewing
-system.  Note that reviewers are not required to look a this material
-when writing their review.
-
-% Acknowledgements should only appear in the accepted version. 
-\section*{Acknowledgments}
-
-\textbf{Do not} include acknowledgements in the initial version of
-the paper submitted for blind review.
-
-If a paper is accepted, the final camera-ready version can (and
-probably should) include acknowledgements. In this case, please
-place such acknowledgements in an unnumbered section at the
-end of the paper. Typically, this will include thanks to reviewers
-who gave useful comments, to colleagues who contributed to the ideas, 
-and to funding agencies and corporate sponsors that provided financial 
-support.
-
-% In the unusual situation where you want a paper to appear in the
-% references without citing it in the main text, use \nocite
-\nocite{langley00}
-
-\bibliography{paper}
-\bibliographystyle{icml2010}
-
 \end{document}
-
-% This document was modified from the file originally made available by
-% Pat Langley and Andrea Danyluk for ICML-2K. This 2010 version was
-% created by Thorsten Joachims & Johannes Fuernkranz, 
-% slightly modified from the 2009 version by Kiri Wagstaff and 
-% Sam Roweis's 2008 version, which is slightly modified from 
-% Prasad Tadepalli's 2007 version which is a lightly 
-% changed version of the previous year's version by Andrew Moore, 
-% which was in turn edited from those of Kristian Kersting and 
-% Codrina Lauth. Alex Smola contributed to the algorithmic style files.

### 2012-03-15 processAndPredict.py
use the method with the highest accuracy

@@ -34,17 +34,26 @@ def checkAccuracy(pred, printFailed=False):
     if printFailed: print failed
     print 'Accuracy: ', len(correct)/(9.*9)
     print 'Accuracy for numbers: ', correctNumbers, '/', 42, '->', correctNumbers/42.
+    return len(correct)/(9.*9)
 
 def prediction(name, value=[]):
+    global bestacc, bestp
     pred = value if len(value) else [int(x[0]) for x in input[name]]
     if -1 in pred: pred = []
     print name, pred
-    if len(pred): checkAccuracy(pred)
+    if len(pred):
+        acc = checkAccuracy(pred)
+        if acc > bestacc:
+            print 'better', bestp, name, bestacc, acc
+            bestacc = acc
+            bestp = name
     print
     return pred
 
 def main():
-    global input, field
+    global input, field, bestacc, bestp
+    bestacc = -1
+    bestp   = 'pNN'
     print "processing input images with octave and writing file octave/inputAndPredictedValues.mat"
     os.system('run-processInput.bat')
 
@@ -60,7 +69,8 @@ def main():
     pLR      = prediction('pLR')
     pPyBrain = prediction('pPyBrain', predict())
 
-    predictMethod = 'pNN'
+    # predictMethod = 'pNN'
+    predictMethod = bestp
     print '-> Using ', predictMethod
     p = eval(predictMethod)
 

### 2012-03-15 doc/icml2010stylefiles/paper.tex
- added grid search for logistic regression
- first draft of paper

@@
 Then we compare different machine learning techniques for supervised learning using the generated data.
-Once the training is complete, images of real Sudokus are captured using the webcam.
-The different methods then try to correctly predict each number.
+Once the training is complete, images of real Sudokus are captured using a webcam and preprocessed to be predicted by the learned models.
+We evaluate the different approaches by analyzing their performance on different data sets.
 \end{abstract}
@@
 This implies that we do not train our models to detect empty fields.
-We could reliably detect empty fields by performing some preprocessing steps on the images (filling borders, adjusting thresholds etc.) thus applying a machine learning technique does not seem appropriate for this task.
+We could detect empty fields with high accuracy by performing some preprocessing steps on the images (filling borders, adjusting thresholds etc.) thus applying a machine learning technique does not seem appropriate for this task.
 
@@
 \subsection{Neural Network}
-Applying a NN seems promising as it was already successfully applied to the quite similar problem of detecting handwritten digits.
-\color{red}ref!\color{black} For the NN we choose a simple architecture with 25 hidden sigmoidal units (+ bias unit), as shown in figure \ref{fig:nn}.
+Applying a NN seems promising as it was already successfully applied to the quite similar problem of detecting handwritten digits.\footnote{http://www.ml-class.org} For the NN we choose a simple architecture with 25 hidden sigmoidal units (+ bias unit), as shown in figure \ref{fig:nn}.
 The size of the input layer is 400 (given by the 20 x 20 pixel images) and the size of the output layer is defined by the number of classes, in this case the numbers 1 to 9.
@@
 Grid search with cross validation resulted in accuracies illustrated in Figure \ref{fig:svm_grid_search}.
-From the best parameter pairs we choose the one that performs best on the validation set, which gives us $C = 20$ and $\gamma = 0.1$.
+We can see that there is a certain area of $(C,\gamma)$ pairs which provide an accuracy close to $100\%$.
+To prevent overfitting we decide to choose $C$ and $\gamma$ by cross checking good pairs against another validation set.
+Further refinement results in the following values $C = 20$ and $\gamma = 0.1$.
 \begin{figure}[ht]
@@
 \begin{itemize}
-\item d1: 4500 generated images
-\item d2: $\sim$ 1000 gathered images
-\item d3: D1 $\cup$ D2
-\item dVal: images of a captured Sudoku
+\item D1: 4500 generated images
+\item D2: $\sim$ 1000 gathered images
+\item D3: D1 $\cup$ D2
+\item DVal: images of some newly captured Sudoku
 \end{itemize}
 
-\color{red}
 \begin{table}[t]
-\caption{Classification accuracies for Neural Network, SVM and Logistic Regression on different data sets.}
+\caption{Classification accuracies for Neural Network, SVM and Logistic Regression on different data sets. All values are rounded averages!}
 \label{sample-table}
@@
 \begin{sc}
-\begin{tabular}{lcccc}
+\begin{tabular}{lccccc}
 \hline
 \abovespace\belowspace
-Data set & NN & SVM & LR & Winner \\
+Data & NN & pyNN & SVM & LR & Winner \\
 \hline
 \abovespace
-D1		& 95.9$\pm$ 0.2& 96.7$\pm$ 0.2& 96.7$\pm$ 0.2& ?\\
-D2		& 83.3$\pm$ 0.6& 80.0$\pm$ 0.6& ? & ?\\
-D3		& 61.9$\pm$ 1.4& 83.8$\pm$ 0.7& ? & ? \\
+D1 		& 78.0 &  & 100 & 80.80 & ?\\
+D2		& 93.4 & 95.94 & 94.41 & 91.36 & pyNN\\
+D3		&  & & & & ? \\
 \belowspace
-DVal	& 74.8$\pm$ 0.5& 78.3$\pm$ 0.6& ? & ?        \\
+DVal	& 89.3 & 88.1 & 92.7 & 85.3 & SVM\\
 \hline
@@
 \end{table}
-\color{black}
+As expected, on data sets D1, D2 and D3 all approaches performed pretty well as they were used for training.
+DVal is the data set that contains completely new images that were never used for training by any of the models.
+Thus, these values are the most interesting ones for evaluation.
+SVM and NN have comparable performance.
+Yet, the SVM is cheaper to train which is why applying an SVM seems more appropriate than using a NN.
+LR is outperformed by the other approaches but predicts with acceptable performance and requires less effort to train.
 
 \section{Future Work}
-\color{red}
-\begin{itemize}
-\item include information if a sudoku is unsolvable
-\item online learning
-\end{itemize}
-\color{black}
-
-\section{References}
+Several ideas arose during this project.
+A first step could be to implement online learning to empower the models to evolute over time and increase their accuracy with more Sudokus being scanned.
+When detecting new Sudokus it turned out that empty field detection was not as reliable as expected, especially when light conditions were bad.
+Improving this part of the system or maybe doing it from a machine learning point of view (adding a new label for empty fields) might be a good idea.
+Another improvement of the systems accuracy could be achieved by including information of the Sudoku solver into field prediction.
+The solver could inform about impossible field combinations forcing the number prediction script to maybe choose a less probable number combination from the model point of view, yet, being a possible one.
 

### 2012-03-15 octave/gridSearchLR.m

@@ -1,0 +1,53 @@
+% Initialization
+%clear ; close all; clc
+
+l = [2^-15, 2^-13, 2^-11, 2^-9, 2^-7, 2^-5, 2^-3, 2^-1, 2^1,  2^3, 2^4, 2^5];
+
+if !exist('config.mat')
+	config();
+endif
+load('config.mat');
+% Load images
+cache = sprintf("%s/%s", data_folder, cache_file);
+if exist(cache)
+    fprintf('Loading generated images (from cache)...\n')
+    load(cache);
+else
+    fprintf('Loading generated images...\n')
+    [X,y] = processImages(data_folder);
+    save("-mat-binary", cache, "X", "y");
+endif
+
+% mix data
+X(:,size(X,2)+1) = y;
+all_indexes_mixed = randperm(size(X,1));
+y = X(all_indexes_mixed, size(X,2));
+X = X(all_indexes_mixed,1:size(X,2)-1);
+
+n_samples = length(y)
+set_size = ceil(length(y)/number_of_sets)
+acc = [];
+
+for lIt=1:length(l)
+	average_accuracy_log_reg = 0;
+	for i = 0:number_of_sets-1
+		printf('\n===== Starting iteration %i (next training/test set combination) =====\n', i);
+		b = min((i+1)*set_size, size(y)); % needed if last set is smaller
+		sel = all_indexes_mixed(i*set_size+1 : b); % selecting the current set of row indexes
+		
+		% define training and test sets
+		X_train = X(complement(sel, 1:size(X,1)),1:size(X,2)-1);
+		y_train = X(complement(sel, 1:size(X,1)),size(X,2));
+		X_val = X(sel, 1:size(X,2)-1);
+		y_val = X(sel, size(X,2));
+		
+		Theta = logRegTrain(X_train,y_train, lIt, maxIter);
+		p = logRegPredict(Theta,X_val);
+		acc = mean(double(p == y_val))*100;
+		average_accuracy_log_reg += length(y_val)*acc;
+	end
+	average_accuracy_log_reg /= size(X,1);
+	acc(length(acc)+1) = average_accuracy_log_reg;
+end
+acc
+save("-mat-binary", "gridSearchLRResult.mat", "l", "acc");

### 2012-03-15 doc/icml2010stylefiles/paper.tex
- fixed paper
- lr gridsearch results added
fix

@@
 \begin{abstract}
-The goal of this work is to implement number recognition for a Soduko solver.
-First we will describe the generation of synthetic data, in this case labelled images of numbers.
+The goal of this work is to implement number recognition for a Sudoku solver.
+First we will describe the generation of synthetic data, in this case labeled images of numbers.
 Then we compare different machine learning techniques for supervised learning using the generated data.
-Once the training is complete, images of real Sudokus are captured using a webcam and preprocessed to be predicted by the learned models.
+Once the training is complete, images of real Sudokus are captured using a webcam and preprocessed in order to be recognized by the learned models.
 We evaluate the different approaches by analyzing their performance on different data sets.
@@
 First of all we generate data by sampling position, rotation, size and color of a number and place it based on these values on a white 20x20 pixel image.
-Examples of 9 such generated images are shown in Figure \ref{fig:generated_numbers}.The corresponding labels are defined by the name of the folder the images are stored in.
+Examples of 9 such generated images are shown in Figure \ref{fig:generated_numbers}.
+The corresponding labels are defined by the name of the folder the images are stored in.
 \begin{figure}[ht]

### 2012-03-15 octave/gridSearchNN.m
added nn grid search and fixed lr grid search

@@ -18,18 +18,20 @@ else
     save("-mat-binary", cache, "X", "y");
 endif
 
+
 % mix data
 X(:,size(X,2)+1) = y;
+
+size(X)
+
 all_indexes_mixed = randperm(size(X,1));
-y = X(all_indexes_mixed, size(X,2));
-X = X(all_indexes_mixed,1:size(X,2)-1);
 
 n_samples = length(y)
 set_size = ceil(length(y)/number_of_sets)
 acc = [];
 
 for lIt=1:length(l)
-	average_accuracy_log_reg = 0;
+	average_accuracy = 0;
 	for i = 0:number_of_sets-1
 		printf('\n===== Starting iteration %i (next training/test set combination) =====\n', i);
 		b = min((i+1)*set_size, size(y)); % needed if last set is smaller
@@ -41,13 +43,13 @@ for lIt=1:length(l)
 		X_val = X(sel, 1:size(X,2)-1);
 		y_val = X(sel, size(X,2));
 		
-		Theta = logRegTrain(X_train,y_train, l(lIt), maxIter);
-		p = logRegPredict(Theta,X_val);
-		acc = mean(double(p == y_val))*100;
-		average_accuracy_log_reg += length(y_val)*acc;
+		[Theta1, Theta2] = nnTrain(X_train, y_train, l(lIt), maxIter);
+		pred = nnPredict(Theta1, Theta2, X_val);
+		acc = mean(double(pred == y_val))*100;
+		average_accuracy += length(y_val)*acc;
 	end
-	average_accuracy_log_reg /= size(X,1);
-	acc(length(acc)+1) = average_accuracy_log_reg;
+	average_accuracy /= size(X,1);
+	acc(length(acc)+1) = average_accuracy;
 end
 acc
-save("-mat-binary", "gridSearchLRResult.mat", "l", "acc");
+save("-mat-binary", "gridSearchNNResult.mat", "l", "acc");

### 2012-03-15 doc/icml2010stylefiles/paper.tex
Final paper and poster

@@
 This implies that we do not train our models to detect empty fields.
-We could detect empty fields with high accuracy by performing some preprocessing steps on the images (filling borders, adjusting thresholds etc.) thus applying a machine learning technique does not seem appropriate for this task.
+We were able to detect empty fields with high accuracy by performing some preprocessing steps on the images (filling borders, adjusting thresholds etc.) thus applying a machine learning technique does not seem appropriate for this task.
+Nevertheless the system can be easily adjusted to handle empty fields as well.
 
@@
 For NNs we test two implementations.
-On the one hand we use a partly self-developed octave code for two-layer NNs which was written as part of the stanford machine learning course\footnote{http://www.ml-class.org}, on the other hand we include PyBrain \cite{schaul2010pybrain} for comparison and easier evaluation of more sophisticated network architectures.
-To find a good value for the regularization parameter $\lambda$ we perform a grid search with k-fold cross-validation which gives us a good performance when choosing $\lambda = 1$.
+On the one hand we use a partly self-developed octave code for two-layer NNs which was written as part of the Stanford machine learning course\footnote{http://www.ml-class.org}, on the other hand we include PyBrain \cite{schaul2010pybrain} for comparison and easier evaluation of more sophisticated network architectures.
+To find a good value for the regularization parameter $\lambda$ we perform a grid search with k-fold cross-validation which gives us a good performance when choosing $\lambda = 1$ for the octave version and $\lambda = 0$ for PyBrain.
 
@@
 \end{align*}
-Using a radial basis function (RBF) as kernel we perform a grid search over $C$ and $\gamma$ to find those parameters that provide highest accuracy on the test set.
+Using a radial basis function (RBF) as a kernel we perform a grid search over $C$ and $\gamma$ to find those parameters that provide highest accuracy on the test set.
 We use the open source library ``libsvm'' \cite{chang2011libsvm} wrapped by some octave scripts for setting up data sets and performing grid search with k-fold cross-validation.
@@
 with respect to $w$.
-Minimization is performed by an algorithm\footnote{http://sprinkler.googlecode.com/svn/trunk/\\regression/fmincg.m} that was provided during the stanford machine learning course.
+Minimization is performed by an algorithm\footnote{http://sprinkler.googlecode.com/svn/trunk/\\regression/fmincg.m} that was provided during the Stanford machine learning course.
 Grid search and k-fold cross-validation recommend a value for the regularization parameter $\lambda$.
@@
 Images captured by the webcam are continously analyzed to find a Sudoku square.
-Images are transformed to greyscale, then an adaptive threshold algorithm is used to preprocess the image for contour detection by OpenCV.
-OpenCV returns a list of contours which are further checked for certain criteria like having four corners or being of certain size.
-Since the contours are not detected precise enough lines are fitted through six checkpoints on each edge in order to achieve subpixel accuracy.
+The images are transformed to greyscale, then an adaptive threshold algorithm is used to preprocess the image for contour detection by OpenCV.
+OpenCV returns a list of contours which are further checked for certain criteria like having four corners or being of a certain size.
+Since the contour detection isn't precise enough, lines are fitted through six checkpoints on each edge in order to achieve subpixel accuracy.
 The content enclosed by the corners is perspectively warped to fit a predefined square size.
 Finally, the thresholded Sudoku field is saved to disk.
-A python script is then used to split the whole field into 9x9 subfields with each subfield being flood filled around the corners for noise reduction.
-The format of the subfields now equals to the one provided by the generated images which allows us to apply the trained NN/SVM/LR models and, finally, apply a Sudoku solver.
+A python script is then used to split the whole field into 9x9 subfields with each subfield being flood filled around the corners for grid-artefact reduction.
+The format of the subfields now equals the one provided by the generated images which allows us to apply the trained NN/SVM/LR models and, finally, apply a Sudoku solver.
 
@@
 \abovespace
-D1 		& 78.0 &  & 100 & 80.80 & ?\\
-D2		& 93.4 & 95.94 & 94.41 & 91.36 & pyNN\\
-D3		&  & & & & ? \\
+D1 		& 86.0 & 84.0 & 98.0 & 83.80 & SVM\\
+D2		& 93.41 & 95.94 & 94.41 & 91.36 & pyNN\\
+D3		& 90.27 & 91.53 & 95.01 & 88.42 & SVM\\
 \belowspace
-DVal	& 89.3 & 88.1 & 92.7 & 85.3 & SVM\\
+DVal	& 87.36 & 86.18 & 91.76 & 83.33 & SVM\\
 \hline

### 2012-03-15 commits.sh
commit history

@@ -1,1 +1,0 @@
-git log | grep -v Date: > commits.txt
